### 2013-04-20 paper.tex
Added initial draft of paper

\documentclass[12pt]{article}
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}

\begin{document}

% Use an empty page style for the title and abstract
\pagestyle{empty}

% =====
% Title
% =====
\title{Hardware-accelerated Rendering of Intersecting Volumes using Boolean Operations}
\author{Andrew Brown \and Joe Geigel}
\date{\small{Rochester Institute of Technology}}
\maketitle

% ========
% Abstract
% ========
\begin{abstract}
Rendering intersecting volumes is difficult because the renderer must alternate
between sampling from each volume to achieve the correct blending.  Methods for
rendering an arbitrary number of intersecting volumes on the GPU have been
developed, but they can require a significant amount of effort to implement.
This paper investigates implementing the simplest case, where two volumes
overlap, using familiar boolean operations and a traditional graphics library.
\end{abstract}

% Switch back to a plain page style for the rest of the paper
\pagestyle{plain}

% ============
% Introduction
% ============
\section{Introduction}

For many years the computing community has created software to visualize data
from medical and scientific equipment.  Normally the data is returned as a
three-dimensional grid of values.  This grid is called a volume, and the
individual values are named voxels, similar to how the elements of a picture are
called pixels.

Volumes produced by such equipment often become quite large, and at first
computers had a difficult time displaying the data in real time.  Early work in
the field thus focused on ways to speed up the process while still maintaining
the highest image quality possible.  In fact, a handful of different methods
with their own unique alterations were tried in the hopes of delivering better
images at a higher frame rate.  In addition, some initial attempts were made at
providing users with better information by simultaneously displaying multiple
volumes in order to place data in its appropriate context.

Of course, as years passed, new technologies developed and it was not long
before high-level, programmable shading languages enabled researchers to utilize
the significant processing power of the main processor on video cards, commonly
called the graphics processing unit, or GPU.  Extremely proficient at performing
similar operations on a large amount of data, GPUs accelerated volume
visualization software tremendously.  Hardware-accelerated implementations of
both the classic slicing method and the high-quality ray-casting method were
developed quickly for single volumes.

With much of the original performance concerns alleviated, and more and more
data being collected from several different sources, researchers are again
looking at ways to effectively display multiple volumes simultaneously.  A few
different methods have been developed for rendering an arbitrary number of
volumes using hardware acceleration, but they can be problematic to implement,
rely on general-purpose GPU programming, or both.

For those reasons, in this paper we examine techniques for rending two volumes
using simple boolean operations.  Although limited, we hope the techniques
presented will be more approachable to software engineers faced with
implementing such technologies, and could help direct future efforts in volume
rendering.

The rest of the paper begins with background information and a more thorough
discussion of previous work in the field.  It then moves on to technologies
available for GPU programming.  Finally, a detailed plan of our proposed work
and its projected implementation is given.

% ==========
% Background
% ==========
\section{Background}

Understanding the techniques presented in this paper is heavily dependent on
knowing some key concepts of how a video card renders three-dimensional objects
by use of a rendering pipeline.  This pipeline is very similar to an assembly
line, where data elements are passed from stage to stage and no element stays in
a stage any longer than it needs to.  It is this unique structure that gives the
video card its speed, since operations are optimized to be performed on a large
amount of data at the same time.

To get the pipeline started, the application feeds vertices to the video card’s
first stage.  So the video card knows which vertices are connected, the
application also provides binding information describing which vertices make up
which polygons.  At this point, the video card’s first stage begins constructing
the polygons and performs some preparatory operations on the vertices, such as
calculating the normal vectors used for illumination and assigning texture
coordinates.  As soon as a polygon is finished, it is immediately passed onto
the next stage, which begins rasterizing the polygons into temporary pixels
called fragments.  The fragments are passed along to the last stage, which
shades the them, decides which ones to discard, and displays those that remain
as pixels.

With the advent of programmable pipelines, developers can now replace some of
the fixed functionality of the pipeline with special programs, called shaders,
that execute custom code, in some cases even making the pipeline do things it
was not designed to do.  The first part of the pipeline that can be replaced is
at the beginning where vertices are prepared for the following stages.  These
shaders are called vertex shaders since they operate on the vertices themselves.
Vertex shaders are normally used to change the position of the vertex, alter its
texture coordinates, or perhaps modify its normal vector.  The second part of
the pipeline that can be replaced is towards the end, where fragments are being
shaded.  Fragment shaders generally change the color of the object or how it is
illuminated.  Many times they look up values stored in textures, which are
generally one, two, or three dimensional images.

Since there are always many fragments generated in rendering an image, designers
have devoted much of the hardware to processing them.  Therefore fragment
shaders are traditionally used to perform the bulk of the work in any
application using a shading language for hardware acceleration.  Because they
are used so much, it is important to note that a fragment’s location is
generally given in normalized texture coordinates.  These coordinates range from
0.0 to 1.0 according to the bounds of the polygon, with 0.0 being one extreme
and 1.0 being the other.

% =============
% Previous Work
% =============
\section{Previous Work}

\subsection{Overview}

The first major breakthrough in volume rendering came with \cite{Levoy88}.
Previous techniques tried to use actual geometry to approximate the surfaces of
volumetric data.  A major problem with these approaches was determining whether
parts of the volume were actually part of a surface or not.  Instead of fitting
geometry to the volume, Levoy’s method samples the volume directly and uses the
resulting vertices to texture a series of polygons aligned with the viewing
direction.  Commonly referred to as slicing, this method was a significant
improvement in accuracy.  Although over twenty years old, sampling the volume
directly has remained the basis for most volume rendering techniques today and
has been termed Direct Volume Rendering.

Levoy made another significant contribution two years later with a paper
\cite{Levoy90} describing Direct Volume Rendering using ray-casting.  Because
many times volumes have large spaces that do not make a difference in the final
image, Levoy focused on having rays skip those spaces by consulting a
hierarchical space division technique known as an octree.  This method has
become known as Empty Space Skipping and is still very important to ray-casting
implementations.

Not long after, Danskin and Hanrahan \cite{Danskin92} improved on Levoy’s work
by integrating another form of acceleration into ray-casting.  Called early ray
termination, the procedure stops the ray if its opacity ever reaches a level
where further samples will not change the final image.  Fairly straightforward
and easy to implement, early ray termination is a staple for most applications.

In 1994 Cullip and Neumann \cite{Cullip94} took another look at the slicing
method.  They primarily compared different methods for orienting the textured
polygons of the slicing method.  The first method oriented the polygons with the
object-space axes.  The other aligned them in image-space.  The former requires
reorienting the polygons as the viewport changes, but the latter must be clipped
to object space requiring a transformation with the inverse viewing matrix.  It
was further suggested that the image-space method would require more passes.
Another important contribution was using a progressive rendering algorithm while
the user is rotating the scene.

Jumping ahead, in \cite{Leu99} Leu and Chen developed a practical platform for
rendering multiple volumes featuring a two-layer approach that separated volume
data from instances of that data.  Consideration for memory management in the
case that all the data did not fit into memory was made, although intersecting
volumes were not considered.

A key source for vital information concerning volume ray-casting with high-level
programmable shading languages came with \cite{Kruger03}.  Kruger demonstrated
that an implementation using fragment shaders and featuring early ray
termination can achieve significant performance gains.  Much of the credit was
given to new features of graphics APIs, including accessing multiple textures,
rendering to arbitrary textures, and replacing depth buffer values.  His
implementation with early-ray termination and ray-casting generally outperformed
other methods by a factor of three and normal ray-casting by a factor of four
for most data sets.

A comprehensive review of the slicing method applied to the GPU can be obtained
from \cite{Ikits04}.  Ikits describes the techniques required for creating the
proxy geometry, texturing the slices using NVIDIA’s Cg shading language, and
using the gradient for local and global illumination models.

\cite{Ruijters06} examined bottlenecks in hardware and devised a two-staged
method for balancing between them.  The first stage breaks the volume into
rectangular bricks, getting rid of empty ones and relieving a bottleneck with
texture memory and the memory bus.  The second stage meanwhile helps in
eliminating details that will not be rendered during rasterization.

Recent work in the field has focused on visualizing multiple volumes.  This
ability helps those that view the data to judge the information in context to
other relevant data.

Grimm \cite{Grimm04} presented techniques to render multiple volumes with the
CPU.

In particular Roßler et.\ al.\ \cite{Rossler06} allowed MRI scans of the human
brain to be viewed simultaneously with the skull by using the slicing method.
An important feature was being able to manipulate each model, in regards to
clipping and opacity, independently of the other.

Later \cite{Rossler08} rendered multiple volumes using the ray-casting method
developed by Kruger.  Rather than render all the volumes at once, which could
sample a large amount of empty space, his technique splits the scene into
multiple regions and generates shaders dynamically.

Some researchers have moved to using general-purpose technologies like CUDA or
OpenCL to better leverage the GPU for multi-volume rendering.  The
implementation by Kainz et. al. \cite{Kainz09} was slower for simple cases, but could
render more complicated scenes.

\subsection{Slicing}

The classic method for performing volume rendering on the GPU is called slicing.
Slicing uses polygons perpendicular to the viewing direction, called proxy
polygons or slices, to sample the volume at regular intervals.  When the entire
volume has been sampled, the slices are composited together to form the final
image.  Because of its relative simplicity, slicing generally yields the fastest
frame rates of all volume rendering methods.

The first step in slicing is to transform the coordinates of the volume’s
bounding box so that they are in terms of the viewing direction.  This is
necessary because the coordinates of the bounding box are stored relative to
itself, but slicing needs the coordinates after any rotations or translations
have been applied.  Luckily, handling this transformation is as simple as
applying the model-view matrix.

Next, the slices are made by generating planes from the closest z position of
the bounding box to the farthest.  The planes are made at regular intervals
according to the image quality desired.  Each resulting plane is tested for
intersections with each edge of the bounding box using the plane-line
intersection test formula.  Any resulting intersection points that fall within
the bounds of the polygon are then sorted into a clockwise or counterclockwise
direction to form the proxy polygon corresponding to that plane.

Now the slices are ready to sample the volume.  A fragment shader is responsible
for performing this operation, looking up voxels based on the texture
coordinates of fragments generated from rendering the slices.  As long as the
vertices of the slices were given texture coordinates that correspond with how
the volume is indexed, the fragment shader’s job is simple.  Except for some
rounding and interpolation, the texture coordinate of the fragment can be used
to access the volume directly, returning a value for that position.  The value
is fed into a transfer function and the resulting color is blended into the
framebuffer.

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{slicing-aligned.pdf}
\caption{Three slices through aligned volume.}
\label{slicing-aligned}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{slicing-rotated.pdf}
\caption{Three slices rotated through volume.}
\label{slicing-rotated}
\end{figure}

Figures \ref{slicing-aligned} and \ref{slicing-rotated} are illustrations
adapted from \cite{Ikits04} showing the process of generating and texturing the
slices.  The first figure shows a trivial example where the volume is aligned
with the viewing direction.  Slices generated are simply rectangles the same
size of the volume.  The volume in the second figure however has been rotated by
the user, and thus the slices are more irregular shapes.

Note that we take the approach of keeping the viewpoint fixed.  Instead of
seeing the viewpoint as rotating around the volume, consider the volume to be
rotating in front of the viewpoint.  We believe this methodology is easier to
understand since the view-space axes are then always aligned with world-space.

\section{Ray Casting}

Ray-casting is an alternate method for performing volume visualization.  Along
with being quite simple and elegant, it is generally considered to produce the
highest-quality images, and has thus become quite popular.  Although usually
slower than slicing, some acceleration methods have been developed that make its
speed quite competitive.

In traditional ray-casting, as in implementations not associated with volume
rendering, rays are projected from the viewpoint through a pixel grid located in
front of the scene.  A ray is created for each pixel in the grid, and each
resulting ray is tested for intersections with every object in the scene.  The
color of the closest intersecting object is then stored as the color for the
corresponding pixel in the grid.  Once all the rays have finished, the values in
the grid are copied to the frame-buffer.  Figure \ref{traditional-ray-casting}
shows an overhead model of the traditional ray-casting process.

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{traditional-ray-casting.pdf}
\caption{Traditional ray-casting with pixel grid.}
\label{traditional-ray-casting}
\end{figure}

To apply ray-casting to volume rendering with a high-level programmable shading
language, \cite{Kruger03} adapted the concept to use fragment shaders on a
volume’s bounding box.  Instead of casting rays through a pixel grid, each
fragment generated by rendering the bounding boxes creates its own ray.  The
rays start at the viewpoint and passes through the fragment’s location.  Any
volumes that intersect the ray are sampled at regular intervals along it by
applying the inverse model view matrix to the points and using them as indices
into the volume.  A transfer function converts the values to colors, and each is
blended into the framebuffer at the fragment’s position.  Figure
\ref{fragment-ray-casting} gives a depiction of the process, with green circles
representing fragments.

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{fragment-ray-casting.pdf}
\caption{Ray-casting by fragments.}
\label{fragment-ray-casting}
\end{figure}

The simplest acceleration technique inherent to ray-casting is early ray
termination, which stops calculation when a certain threshold has been reached
while marching along the ray.  Fairly intuitive, early ray termination works
because if the data has already added up to be opaque, values behind it no
longer make a difference.

Early ray termination is not the only way to accelerate ray-casting. A more
complicated technique is called empty space skipping.  Pioneered by Marc Levoy
in 1990 \cite{Levoy90}, empty space skipping is traditionally done with an
octree.

An octree is a tree data structure used to divide a three-dimensional space,
with each node representing a section of that space.  The first node in the
tree, the root, is the size of the entire space, and as the name implies, it is
broken up into eight smaller sections.  These sections are added as children to
the root node, and then they are split into eight smaller sections themselves.
The process continues until the sections reach a size that the application is
comfortable working with.

With the sizes set, the nodes at the bottom of the tree, called leaf nodes, set
various characteristics based on the objects inside their section of the space.
Nodes in the rest of the tree then use the characteristics of their eight
children nodes to set their own characteristics.  In this way an octree holds
information about the space on many different levels of detail, which an
application can use to intelligently navigate and make decisions about the
space.  Figure \ref{octree} shows the first three levels of an octree.

\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{octree.pdf}
\caption{The first three levels of an octree.}
\label{octree}
\end{figure}

For volume rendering, an octree represents the space of the volume, and its
nodes store whether that section of the space is empty or not.  A leaf node,
which has a width equal to twice the distance between two voxels, is determined
to be empty if all of the eight voxels at its corners are below a certain
threshold.  Larger nodes meanwhile are empty if the eight nodes below them are
all empty.  With an octree of this sort in place, rays can potentially save
themselves a lot of work by checking the nodes in the octree before sampling,
starting with the root and working its way down.  If the ray ever encounters an
empty section, it can skip over all the points along the ray that would fall in
that section.

% ==========
% Approaches
% ==========
\section{Approaches}

Two different approaches.

% Boolean AND

The first approach uses a Boolean AND operation combined with depth-buffer
masking.  It is mostly GPU-based, only requiring that the Boolean operation be
performed on the CPU.  While there is relatively less work on the CPU, it does
require a significant amount of passes.  The pseudocode for the approach is
shown below.  Note that early Z-kill should prevent fragments that are masked
off from being processed by the fragment shader.

\begin{enumerate}
\item Render fragments behind the intersection into a framebuffer
\item Render intersection into another framebuffer
\item Render fragments in front of the intersection to the screen
\item Mask off depth buffer with depths of intersection's front faces
\item Render everything else, sorted by depth, to the screen
\end{enumerate}

% Boolean XOR

The second approach uses a Boolean AND operation together with a Boolean XOR.
It is mostly CPU-based, since it requires both Boolean operations to be done on
the CPU.  The extra work on the CPU results in less rendering passes, however.
The pseudocode for the approach is shown below.

\begin{enumerate}
\item Render pieces from Boolean XOR behind intersection to framebuffer
\item Render intersection using Boolean AND to framebuffer
\item Render pieces from Boolean XOR in front of intersection to screen
\end{enumerate}

% ==============
% Implementation
% ==============
\section{Implementation}

The two approaches were implemented using OpenGL with C++.  To ensure results
are relevant to current and future applications, we used only the core profile
of OpenGL 3.  In other words, no deprecated functionality from OpenGL 2 was
used.

To facilitate the process, a handful of general-purpose graphics libraries were
developed.  At the lowest levels are a simple 3D mathematics library and a
library providing thin C++ wrappers around the components already in OpenGL.
Above that, another library provides additional utilities for working with
OpenGL, including one for loading volumetric data as a texture.  At the highest
level is a library providing a simple render graph that can be used to rapidly
prototype OpenGL applications.  Similar to SVG, the render graph can be written
in XML so scenes can easily be saved, reloaded, and shared.  All of the
libraries can be compiled as shared libraries using GCC on Linux and Mac OS X.

From there an application was created to view the resulting scenes.  It can
rotate the camera around the scene, as well as select objects in the scene and
drag them around.  It shows the current frame rate so we can evaluate the
performance of the scenes.  Some scene graph components that are not general
enough to be in the library were created at the application level.

% ==========
% Evaluation
% ==========
\section{Evaluation}

How fast they can be rendered?
How easy they were to implement?

% ============
% Bibliography
% ============

\bibliographystyle{alpha}
\bibliography{paper.bib}

\end{document}

### 2013-04-21 paper.tex
Added subsections for different approaches
Better introduction to approaches section

@@
-\documentclass[12pt]{article}
+\documentclass{article}
 \setlength{\pdfpagewidth}{8.5in}
@@
 \usepackage[utf8]{inputenc}
+\usepackage{fancyvrb}
 
@@
 
-Two different approaches.
+We take two different approaches to rendering intersecting volumes using boolean
+operations.  The boolean operations involved are AND and XOR.  AND takes the
+intersection of two shapes, while XOR, meaning exclusive OR, takes everything
+but the intersection.
 
-% Boolean AND
+\subsection{Boolean AND with Depth Buffer Masking}
 
@@
 
-% Boolean XOR
+\subsection{Boolean AND with Boolean XOR}
 

### 2013-04-21 paper.tex
Moved note about early z-kill in first approach below pseudocode
Broke up introduction to first approach
Introduction for first approach summarizes the approach
Use default font size
Use the 'fancyvrb' package for better verbatim environments
Fully detailed first approach

@@
 The first approach uses a Boolean AND operation combined with depth-buffer
-masking.  It is mostly GPU-based, only requiring that the Boolean operation be
-performed on the CPU.  While there is relatively less work on the CPU, it does
-require a significant amount of passes.  The pseudocode for the approach is
-shown below.  Note that early Z-kill should prevent fragments that are masked
-off from being processed by the fragment shader.
+masking.  The basic idea is to mask off everything but the area on the screen
+with the intersection, rendering the cubes behind it, the intersection itself,
+and then the cubes in front of it.  Then the area of the intersection is masked
+off and everything else is drawn depth-sorted.
+
+It is mostly GPU-based, only requiring that the Boolean operation be performed
+on the CPU.  While there is relatively less work on the CPU, it does require a
+significant amount of passes.
+
+The pseudocode for the approach is shown below.
 
@@
 \end{enumerate}
+
+The first step in this approach is to bind all the textures.  There are two for
+the volumes, two for storing the texture coordinates of the back faces of the
+cubes, and one for storing results between passes.  The last three textures will
+be used with external framebuffers, so they should have a width and height at
+least as big as the window.
+
+\begin{Verbatim}[fontsize=\small]
+  <texture id="first-volume-texture" file="first.vlb">
+  <texture id="second-volume-texture" file="second.vlb">
+  <texture id="primary-coords-texture" width="768" height="768">
+  <texture id="secondary-coords-texture" width="768" height="768">
+  <texture id="results-texture" width="768" height="768">
+\end{Verbatim}
+
+Next the volumes behind the intersection are rendered into the results texture.
+Because we want the part of the volume behind the intersection we first render
+the texture coordinates of the back faces of the cubes into a framebuffer, then
+render the back faces of the intersection as per \cite{Kruger03}.  Note in the code
+snippet below we only show the first volume, hide where we set the uniform
+variables in the shader program for brevity, and give the intersection an ID so
+it can be redrawn later.
+
+\begin{Verbatim}[fontsize=\small]
+  <cull faces="front" />
+  <framebuffer>
+    <attachment usage="color" link="primary-coords-texture" />
+    <use program="primary-coords-program">
+      <instance of="cube-1-group" />
+    </use>
+  </framebuffer>
+  <framebuffer>
+    <attachment usage="color" link="results-texture" />
+    <use program="first-pass-program">
+      <booleanAnd id="intersection" of="cube-1 cube-2" />
+    </use>
+  </framebuffer>
+\end{Verbatim}
+
+From there the intersection itself is rendered.  Because the intersection needs
+to alternate between samples from both volumes, we have to render the back face
+coordinates into two different textures, and then make them both available to
+the shader program that renders the intersection.
+
+\begin{Verbatim}[fontsize=\small]
+  <cull faces="front" />
+  <framebuffer>
+    <attachment usage="color" link="coords-texture-1" />
+    <use program="primary-coords-program">
+      <instance of="intersection" />
+    </use>
+  </framebuffer>
+  <framebuffer>
+    <attachment usage="color" link="coords-texture-2" />
+    <use program="secondary-coords-program">
+      <instance of="intersection" />
+    </use>
+  </framebuffer>
+  <cull faces="back" />
+  <framebuffer>
+    <attachment usage="color" link="third-pass-texture" />
+    <use program="third-pass-program">
+      <uniform name="CoordsTexture1" link="coords-texture-1" />
+      <uniform name="CoordsTexture2" link="coords-texture-2" />
+      <instance of="intersection" />
+    </use>
+  </framebuffer>
+\end{Verbatim}
+
+After that the parts of the volumes in front of the intersection are rendered to
+the results texture.  We have the shader program set the depth of the fragments
+generated to the lowest value to effectively mask off that part of the screen.
+
+\begin{Verbatim}[fontsize=\small]
+  <cull mode="back" />
+  <framebuffer>
+    <attachment usage="color" link="coords-texture-1" />
+    <use program="primary-coords-program">
+      <uniform name="MVPMatrix" usage="modelviewprojection" />
+      <instance of="cube-1" />
+    </use>
+  </framebuffer>
+  <framebuffer>
+    <attachment usage="color" link="results-texture" />
+    <use program="fourth-pass-program">
+      <instance of="intersection" />
+    </use>
+  </framebuffer>
+\end{Verbatim}
+
+Finally we render the two volumes again depth-sorted with blending enabled and
+the depth function set to \emph{LESS}.  Since on the previous step we masked off
+the area of the screen with the intersection, the fragments in those parts will
+just be discarded.  In fact, on most modern video cards they will not even be
+processed by the fragment shader at all because of early Z-kill.
+
+\begin{Verbatim}[fontsize=\small]
+  <blend enabled="true" />
+  <depth function="less" />
+  <sort>
+    <group>
+      <framebuffer>
+        <attachment usage="color" link="primary-coords-texture" />
+        <use program="primary-coords-program">
+          <cull mode="front" />
+          <instance of="cube-1-group" />
+        </use>
+      </framebuffer>
+      <use program="kruger-program">
+        <cull faces="back" />
+        <group id="cube-1-group">
+          <translate x="-1">
+            <cube id="cube-1" />
+          </translate>
+        </group>
+      </use>
+    </group>
+    <group>
+      <framebuffer>
+        <attachment usage="color" link="coords-texture-2" />
+        <use program="primary-coords-program">
+          <cull mode="front" />
+          <instance of="cube-2-group" />
+        </use>
+      </framebuffer>
+      <use program="kruger-program">
+        <cull mode="back" />
+        <group id="cube-2-group">
+          <translate x="+1">
+            <cube id="cube-2" />
+          </translate>
+        </group>
+      </use>
+    </group>
+  </sort>
+\end{Verbatim}
+
+Note that early Z-kill should prevent fragments that are masked off from being
+processed by the fragment shader.
 

### 2013-04-21 paper.tex
Removed old pseudocode for first approach
Removed old note in first approach about early Z-kill

@@
 
-The pseudocode for the approach is shown below.
-
-\begin{enumerate}
-\item Render fragments behind the intersection into a framebuffer
-\item Render intersection into another framebuffer
-\item Render fragments in front of the intersection to the screen
-\item Mask off depth buffer with depths of intersection's front faces
-\item Render everything else, sorted by depth, to the screen
-\end{enumerate}
-
 The first step in this approach is to bind all the textures.  There are two for
@@
 
-Note that early Z-kill should prevent fragments that are masked off from being
-processed by the fragment shader.
-
 \subsection{Boolean AND with Boolean XOR}

### 2013-04-21 paper.tex
Broke up introduction in second approach
Moved notes about characteristics of second approach to end of subsection
Summary of second approach in its introduction
Fully-detailed second approach

@@
 The second approach uses a Boolean AND operation together with a Boolean XOR.
-It is mostly CPU-based, since it requires both Boolean operations to be done on
-the CPU.  The extra work on the CPU results in less rendering passes, however.
-The pseudocode for the approach is shown below.
+Here the idea is to break up both volumes into pieces around the intersection.
+All the pieces behind the intersection are rendered first, then the intersection
+itself, and then all the pieces in front of the intersection.
 
@@
 \end{enumerate}
+
+Like the first approach, the first step is to set up the textures.  There are
+two textures for the volumes, two to store the texture coordinates of the back
+faces, and one to hold results between passes.  The textures to store the
+texture coordinates and results between passes will be used with external
+framebuffer objects, and should thus be at least as big as the window.
+
+\begin{Verbatim}[fontsize=\small]
+  <texture id="volume-texture-1" file="volume-1.vlb" />
+  <texture id="volume-texture-2" file="volume-2.vlb" />
+  <texture id="coords-texture-1" size="768" />
+  <texture id="coords-texture-2" size="768" />
+  <texture id="results-texture" size="768" />
+\end{Verbatim}
+
+Storage should also be created to act as a depth buffer for some passes.  Since
+we do not need to read from it, it can just be a \emph{renderbuffer}.  It also
+needs to be at least as big as the window.
+
+\begin{Verbatim}[fontsize=\small]
+  <renderbuffer id="depth-renderbuffer" format="depth" size="768" />
+\end{Verbatim}
+
+Next the results texture should be cleared.
+
+\begin{Verbatim}[fontsize=\small]
+  <framebuffer>
+    <attachment usage="color" link="results-texture" />
+    <clear color="0 0 0 0" />
+  </framebuffer>
+\end{Verbatim}
+
+Then the pieces behind the intersection should be used to render the volumes to
+the results texture.  Because there may be multiple pieces behind the
+intersection and we only want the texture coordinates of the faces that are
+farthest back, we first clear the depth buffer to zero and set the depth
+function to \emph{GREATER}.  Once we have those coordinates stored, we do the
+opposite by clearing the depth buffer to one and setting the depth function back
+to \emph{LESS}.  That way we render using the faces that are farthest forward.
+Note in the code snippet below we only show the first volume.
+
+\begin{Verbatim}[fontsize=\small]
+  <framebuffer>
+    <attachment usage="color" link="coords-texture-1" />
+    <attachment usage="depth" link="depth-renderbuffer" />
+    <clear depth="0.0" />
+    <depth function="greater" />
+    <cull faces="front" />
+    <use program="primary-coords-program">
+      <group id="pieces-of-cube-1-behind-intersection-group">
+        <booleanXor of="cube-1 cube-2" hide="cube-2" filter="gequal" />
+      </group>
+    </use>
+  </framebuffer>
+  <framebuffer>
+    <attachment usage="color" link="results-texture" />
+    <attachment usage="depth" link="depth-renderbuffer" />
+    <clear depth="1.0" />
+    <depth function="less" />
+    <cull faces="back" />
+    <use program="first-pass-program">
+      <uniform name="MVPMatrix" usage="viewprojection" />
+      <uniform name="BackFacesTexture" link="coords-texture-1" />
+      <uniform name="VolumeTexture" link="volume-texture-1" />
+      <instance of="pieces-of-cube-1-behind-intersection-group" />
+    </use>
+  </framebuffer>
+\end{Verbatim}
+
+Following our pattern, we now render the intersection itself.  Notice we do not
+need a depth buffer for this pass, although we do need to make sure both sets of
+texture coordinates are stored and are available when we render the
+intersection.
+
+\begin{Verbatim}[fontsize=\small]
+  <cull faces="front" />
+  <framebuffer>
+    <attachment usage="color" link="coords-texture-1" />
+    <use program="primary-coords-program">
+      <uniform name="MVPMatrix" usage="viewprojection" />
+      <group id="intersection-group">
+        <booleanAnd of="cube-1 cube-2" />
+      </group>
+    </use>
+  </framebuffer>
+  <framebuffer>
+    <attachment usage="color" link="coords-texture-2" />
+    <use program="secondary-coords-program">
+      <uniform name="MVPMatrix" usage="viewprojection" />
+      <instance of="intersection-group" />
+    </use>
+  </framebuffer>
+  <cull faces="back" />
+  <framebuffer>
+    <attachment usage="color" link="results-texture" />
+    <use program="second-pass-program">
+      <uniform name="MVPMatrix" usage="viewprojection" />
+      <uniform name="ResultsTexture" link="results-texture" />
+      <uniform name="CoordsTexture1" link="coords-texture-1" />
+      <uniform name="CoordsTexture2" link="coords-texture-2" />
+      <uniofrm name="VolumeTexture1" link="volume-texture-1" />
+      <uniofrm name="VolumeTexture2" link="volume-texture-2" />
+      <instance of="intersection-group" />
+    </use>
+  </framebuffer>
+\end{Verbatim}
+
+Next we render any pieces of the first cube in front of the intersection to the
+results texture.  Similar to the first pass, we use the depth buffer to make
+sure we only take the texture coordinates of the faces we want.  Again we omit
+the code to render the second cube.
+
+\begin{Verbatim}[fontsize=\small]
+  <framebuffer>
+    <attachment usage="color" link="coords-texture-1" />
+    <attachment usage="depth" link="depth-renderbuffer" />
+    <cull faces="front" />
+    <depth function="greater" />
+    <clear depth="0.0" />
+    <use program="primary-coords-program">
+      <uniform name="MVPMatrix" usage="viewprojection" />
+      <group id="pieces-of-cube-1-in-front-of-intersection">
+        <booleanXor of="cube-1 cube-2" hide="cube-2" filter="less" />
+      </group>
+    </use>
+  </framebuffer>
+  <framebuffer>
+    <attachment usage="color" link="results-texture" />
+    <attachment usage="depth" link="depth-renderbuffer" />
+    <depth function="less" />
+    <cull faces="back" />
+    <use program="third-pass-program">
+      <uniform name="MVPMatrix" usage="viewprojection" />
+      <uniform name="ResultsTexture" link="results-texture" />
+      <uniform name="CoordsTexture" link="coords-texture-1" />
+      <uniform name="VolumeTexture" link="volume-texture-1" />
+      <insance of="pieces-of-cube-1-in-front-of-intersection" />
+    </use>
+  </framebuffer>
+\end{Verbatim}
+
+Finally we render the results to the screen by drawing both cubes again.  The
+shader program for this final pass just outputs whatever is in the results
+texture.  We need this final pass because depending on how the scene is laid out
+there may not be any pieces behind or in front of the intersection, meaning the
+first and third passes may not be run.
+
+\begin{Verbatim}[fontsize=\small]
+  <use program="final-pass-program">
+    <uniform name="ResultsTexture" link="results-texture" />
+    <translate x="-1.0">
+      <uniform name="MVPMatrix" usage="modelviewprojection" />
+      <cube id="cube-1" />
+    </translate>
+    <translate x="+1.0">
+      <uniform name="MVPMatrix" usage="modelviewprojection" />
+      <cube id="cube-2" />
+    </translate>
+  </use>
+\end{Verbatim}
+
+It is mostly CPU-based, since it requires both Boolean operations to be done on
+the CPU.  The extra work on the CPU results in less rendering passes, however.
+The pseudocode for the approach is shown below.
 

### 2013-04-21 paper.tex
Removed old pseudocode for second approach
Included boolean diagrams in LaTeX file

@@
 
+\begin{figure}
+\centering
+\includegraphics[width=0.8\textwidth]{boolean-and.pdf}
+\caption{
+Top down view of rendering intersecting volumes using the Boolean AND with depth
+buffer masking approach.  First the part of the volume directly behind the
+intersection, colored red, is rendered.  Next the intersection, colored green,
+is rendered.  Then the the part of the volume directly in front of the
+intersection, colored blue, is rendered.  Finally the depth buffer is masked off
+using the intersection, and everything else, colored yellow, is rendered.
+}
+\label{boolean-and}
+\end{figure}
+
 It is mostly GPU-based, only requiring that the Boolean operation be performed
@@
 
-\begin{enumerate}
-\item Render pieces from Boolean XOR behind intersection to framebuffer
-\item Render intersection using Boolean AND to framebuffer
-\item Render pieces from Boolean XOR in front of intersection to screen
-\end{enumerate}
+\begin{figure}
+\centering
+\includegraphics[width=0.8\textwidth]{boolean-xor.pdf}
+\caption{
+Rendering intersecting volumes using the Boolean AND with Boolean XOR approach.
+First the pieces behind the intersection, colored red, are rendered.  Next the
+intersection, colored green, is rendered.  Then the pieces in front of the
+intersection, colored blue, are rendered.  A final pass renders the results to
+the screen.
+}
+\label{boolean-xor}
+\end{figure}
 

### 2013-04-22 paper.tex
Put bibliography on a new page
Better introduction to first approach

@@
 The first approach uses a Boolean AND operation combined with depth-buffer
-masking.  The basic idea is to mask off everything but the area on the screen
-with the intersection, rendering the cubes behind it, the intersection itself,
-and then the cubes in front of it.  Then the area of the intersection is masked
-off and everything else is drawn depth-sorted.
+masking.  The basic idea is render the part of the volume directly behind the
+intersection, render the intersection itself, and then render the part of the
+volume directly in front of the intersection.  Finally the area of the
+intersection is masked off and everything else is drawn depth-sorted with
+blending enabled.
 
@@
 
+\newpage
 \bibliographystyle{alpha}

### 2013-04-22 paper.tex
Split up introduction to approaches
Briefly go over Kruger method again in introduction to approaches

@@
 We take two different approaches to rendering intersecting volumes using boolean
-operations.  The boolean operations involved are AND and XOR.  AND takes the
-intersection of two shapes, while XOR, meaning exclusive OR, takes everything
-but the intersection.
+operations.
+
+The boolean operations involved are AND and XOR.  AND takes the intersection of
+two shapes, while XOR, meaning exclusive OR, takes everything but the
+intersection.
+
+Both approaches are based on the volume rendering method in \cite{Kruger03}, as
+previously mentioned.  To render part of a volume, they first render the texture
+coordinates of the back faces of that part to a texture.  Then the front faces
+of that part are rendered with a shader that uses the previously-stored texture
+coordinates to form a ray through the volume.  The ray then samples the volume
+at regular intervals.
 

### 2013-04-22 paper.tex
Mention that in both approaches passes must start with results of previous pass
Fixed using 'primary-coords-texture' and 'secondary-coords-texture' in some places

@@
 
+In addition, both approaches use multiple passes.  To blend the samples
+correctly between passes, each pass must start with the results of the previous
+pass.  The results of all the passes cannot be simply blended together
+afterwards.
+
 \subsection{Boolean AND with Depth Buffer Masking}
@@
   <texture id="second-volume-texture" file="second.vlb">
-  <texture id="primary-coords-texture" width="768" height="768">
-  <texture id="secondary-coords-texture" width="768" height="768">
+  <texture id="coords-texture-1" width="768" height="768">
+  <texture id="coords-texture-2" width="768" height="768">
   <texture id="results-texture" width="768" height="768">
@@
   <framebuffer>
-    <attachment usage="color" link="primary-coords-texture" />
+    <attachment usage="color" link="coords-texture-1" />
     <use program="primary-coords-program">
@@
       <framebuffer>
-        <attachment usage="color" link="primary-coords-texture" />
+        <attachment usage="color" link="coords-texture-1" />
         <use program="primary-coords-program">

### 2013-04-22 paper.tex
Fixed instancing booleanAnd directly in first approach

@@
     <use program="first-pass-program">
-      <booleanAnd id="intersection" of="cube-1 cube-2" />
+      <group id="intersection-group">
+        <booleanAnd of="cube-1 cube-2" />
+      </group>
     </use>
@@
     <use program="primary-coords-program">
-      <instance of="intersection" />
+      <instance of="intersection-group" />
     </use>
@@
     <use program="secondary-coords-program">
-      <instance of="intersection" />
+      <instance of="intersection-group" />
     </use>
@@
       <uniform name="CoordsTexture2" link="coords-texture-2" />
-      <instance of="intersection" />
+      <instance of="intersection-group" />
     </use>
@@
     <use program="fourth-pass-program">
-      <instance of="intersection" />
+      <instance of="intersection-group" />
     </use>

### 2013-04-22 paper.tex
Fixed instancing first cube directly in first approach

@@
       <uniform name="MVPMatrix" usage="modelviewprojection" />
-      <instance of="cube-1" />
+      <instance of="cube-1-group" />
     </use>

### 2013-04-22 paper.tex
Replaced XML for first approach with pseudocode

--- 2013-04-22 paper.tex
@@
 intersection, colored blue, is rendered.  Finally the depth buffer is masked off
-using the intersection, and everything else, colored yellow, is rendered.
+using the Boolean AND so that when the cubes are rendered again, only the parts
+colored yellow are rendered.
 }
@@
 
-\begin{Verbatim}[fontsize=\small]
-  <texture id="first-volume-texture" file="first.vlb">
-  <texture id="second-volume-texture" file="second.vlb">
-  <texture id="coords-texture-1" width="768" height="768">
-  <texture id="coords-texture-2" width="768" height="768">
-  <texture id="results-texture" width="768" height="768">
-\end{Verbatim}
+\begin{enumerate}
+  \item Bind volume textures
+  \item Bind textures to store texture coordinates
+  \item Bind texture to store results
+\end{enumerate}
 
-Next the volumes behind the intersection are rendered into the results texture.
-Because we want the part of the volume behind the intersection we first render
-the texture coordinates of the back faces of the cubes into a framebuffer, then
-render the back faces of the intersection as per \cite{Kruger03}.  Note in the code
-snippet below we only show the first volume, hide where we set the uniform
-variables in the shader program for brevity, and give the intersection an ID so
-it can be redrawn later.
+Next the part of the volume behind the intersection is rendered into the results
+texture.  To get just that part the back faces of the Boolean AND are used
+instead of the front faces of the cube.
 
-\begin{Verbatim}[fontsize=\small]
-  <cull faces="front" />
-  <framebuffer>
-    <attachment usage="color" link="coords-texture-1" />
-    <use program="primary-coords-program">
-      <instance of="cube-1-group" />
-    </use>
-  </framebuffer>
-  <framebuffer>
-    <attachment usage="color" link="results-texture" />
-    <use program="first-pass-program">
-      <booleanAnd id="intersection" of="cube-1 cube-2" />
-    </use>
-  </framebuffer>
-\end{Verbatim}
+\begin{enumerate}
+  \item Cull front faces
+  \item Bind framebuffer with coordinates texture
+  \item Activate shader program for rendering texture coordinates
+  \item Draw cube behind intersection
+  \item Bind framebuffer with results texture
+  \item Activate shader program for rendering a volume
+  \item Draw Boolean AND of cubes
+\end{enumerate}
 
@@
 
-\begin{Verbatim}[fontsize=\small]
-  <cull faces="front" />
-  <framebuffer>
-    <attachment usage="color" link="coords-texture-1" />
-    <use program="primary-coords-program">
-      <instance of="intersection" />
-    </use>
-  </framebuffer>
-  <framebuffer>
-    <attachment usage="color" link="coords-texture-2" />
-    <use program="secondary-coords-program">
-      <instance of="intersection" />
-    </use>
-  </framebuffer>
-  <cull faces="back" />
-  <framebuffer>
-    <attachment usage="color" link="third-pass-texture" />
-    <use program="third-pass-program">
-      <uniform name="CoordsTexture1" link="coords-texture-1" />
-      <uniform name="CoordsTexture2" link="coords-texture-2" />
-      <instance of="intersection" />
-    </use>
-  </framebuffer>
-\end{Verbatim}
+\begin{enumerate}
+  \item Cull front faces
+  \item Bind framebuffer with first coordinates texture
+  \item Activate shader program for rendering primary texture coordinates
+  \item Draw Boolean AND of cubes
+  \item Bind framebuffer with second coordinates texture
+  \item Activate shader program for rendering secondary texture coordinates
+  \item Draw Boolean AND of cubes
+  \item Cull back faces
+  \item Bind framebuffer with results texture
+  \item Activate shader program for rendering volume
+  \item Draw Boolean AND of cubes
+\end{enumerate}
 
-After that the parts of the volumes in front of the intersection are rendered to
-the results texture.  We have the shader program set the depth of the fragments
-generated to the lowest value to effectively mask off that part of the screen.
+After that the part of the volume in front of the intersection is rendered to
+the results texture.
 
-\begin{Verbatim}[fontsize=\small]
-  <cull mode="back" />
-  <framebuffer>
-    <attachment usage="color" link="coords-texture-1" />
-    <use program="primary-coords-program">
-      <uniform name="MVPMatrix" usage="modelviewprojection" />
-      <instance of="cube-1" />
-    </use>
-  </framebuffer>
-  <framebuffer>
-    <attachment usage="color" link="results-texture" />
-    <use program="fourth-pass-program">
-      <instance of="intersection" />
-    </use>
-  </framebuffer>
-\end{Verbatim}
+\begin{enumerate}
+  \item Cull back faces
+  \item Bind framebuffer with coordinates texture
+  \item Activate shader program for rendering primary texture coordinates
+  \item Draw cube in front of intersection
+  \item Bind framebuffer with results texture
+  \item Activate shader program for rendering volume
+  \item Draw Boolean AND of cubes
+  \item Mask off depth buffer
+\end{enumerate}
+
+The area of the screen with the intersection should be masked off so that
+nothing else in that area will be rendered.
+
+\begin{enumerate}
+  \item Activate shader program for setting depth to zero
+  \item Draw Boolean AND of cubes
+\end{enumerate}
 
@@
 
-\begin{Verbatim}[fontsize=\small]
-  <blend enabled="true" />
-  <depth function="less" />
-  <sort>
-    <group>
-      <framebuffer>
-        <attachment usage="color" link="coords-texture-1" />
-        <use program="primary-coords-program">
-          <cull mode="front" />
-          <instance of="cube-1-group" />
-        </use>
-      </framebuffer>
-      <use program="kruger-program">
-        <cull faces="back" />
-        <group id="cube-1-group">
-          <translate x="-1">
-            <cube id="cube-1" />
-          </translate>
-        </group>
-      </use>
-    </group>
-    <group>
-      <framebuffer>
-        <attachment usage="color" link="coords-texture-2" />
-        <use program="primary-coords-program">
-          <cull mode="front" />
-          <instance of="cube-2-group" />
-        </use>
-      </framebuffer>
-      <use program="kruger-program">
-        <cull mode="back" />
-        <group id="cube-2-group">
-          <translate x="+1">
-            <cube id="cube-2" />
-          </translate>
-        </group>
-      </use>
-    </group>
-  </sort>
-\end{Verbatim}
+\begin{enumerate}
+  \item Enable blending
+  \item Set depth function to \emph{LESS}
+  \item Bind framebuffer with coordinates texture
+  \item Cull front faces
+  \item Activate shader program for rendering primary texture coordinates
+  \item Draw cube behind intersection
+  \item Cull back faces
+  \item Activate shader program for rendering volumes
+  \item Draw cube behind intersection
+  \item Repeat for cube in front of intersection
+\end{enumerate}
 
@@
 
-Next we render any pieces of the first cube in front of the intersection to the
+Next we render any pieces of the cube in front of the intersection to the
 results texture.  Similar to the first pass, we use the depth buffer to make

### 2013-04-22 paper.tex
Simplified explanation of rendering part behind intersection in first approach

--- 2013-04-22 paper.tex
@@
 
-Next the volumes behind the intersection are rendered into the results texture.
-Because we want the part of the volume behind the intersection we first render
-the texture coordinates of the back faces of the cubes into a framebuffer, then
-render the back faces of the intersection as per \cite{Kruger03}.  Note in the code
-snippet below we only show the first volume, hide where we set the uniform
-variables in the shader program for brevity, and give the intersection an ID so
-it can be redrawn later.
+Next the part of the volume behind the intersection is rendered into the results
+texture.  To get just that part the back faces of the Boolean AND are used
+instead of the front faces of the cube.  Note in the code snippet below we only
+show the first volume, hide where we set the uniform variables in the shader
+program for brevity, and give the intersection an ID so it can be redrawn later.
 

### 2013-04-22 paper.tex
Use 'omit' in explanation of rendering volume behind intersection in first approach

@@
 texture.  To get just that part the back faces of the Boolean AND are used
-instead of the front faces of the cube.  Note in the code snippet below we only
-show the first volume, hide where we set the uniform variables in the shader
-program for brevity, and give the intersection an ID so it can be redrawn later.
+instead of the front faces of the cube.  Note in the code snippet below we omit
+the instructions for rendering the second volume and for setting uniform
+variables.  We also give the intersection an ID so it can be redrawn later.
 

### 2013-04-22 paper.tex
Fixed not mentioning group in explanation of rendering volume behind intersection in first approach
Only one part of volume in front of intersection in first approach

@@
 the instructions for rendering the second volume and for setting uniform
-variables.  We also give the intersection an ID so it can be redrawn later.
+variables.  We also put the Boolean AND into a group so it can be redrawn later
+without duplicating the calculations or geometry.
 
@@
 
-After that the parts of the volumes in front of the intersection are rendered to
+After that the part of the volume in front of the intersection is rendered to
 the results texture.  We have the shader program set the depth of the fragments

### 2013-04-22 paper.tex
Fixed using 'mode' attribute for 'cull' node instead of 'faces'

@@
 \begin{Verbatim}[fontsize=\small]
-  <cull mode="back" />
+  <cull faces="back" />
   <framebuffer>
@@
         <use program="primary-coords-program">
-          <cull mode="front" />
+          <cull faces="front" />
           <instance of="cube-1-group" />
@@
         <use program="primary-coords-program">
-          <cull mode="front" />
+          <cull faces="front" />
           <instance of="cube-2-group" />
@@
       <use program="kruger-program">
-        <cull mode="back" />
+        <cull faces="back" />
         <group id="cube-2-group">

### 2013-04-22 paper.tex
Replaced XML for second approach with pseudocode

--- 2013-04-22 paper.tex
@@
 
-\begin{Verbatim}[fontsize=\small]
-  <texture id="first-volume-texture" file="first.vlb">
-  <texture id="second-volume-texture" file="second.vlb">
-  <texture id="coords-texture-1" width="768" height="768">
-  <texture id="coords-texture-2" width="768" height="768">
-  <texture id="results-texture" width="768" height="768">
-\end{Verbatim}
+\begin{enumerate}
+  \item Bind volume textures
+  \item Bind textures to store texture coordinates
+  \item Bind texture to store results
+\end{enumerate}
 
@@
 texture.  To get just that part the back faces of the Boolean AND are used
-instead of the front faces of the cube.  Note in the code snippet below we omit
-the instructions for rendering the second volume and for setting uniform
-variables.  We also put the Boolean AND into a group so it can be redrawn later
-without duplicating the calculations or geometry.
+instead of the front faces of the cube.
 
-\begin{Verbatim}[fontsize=\small]
-  <cull faces="front" />
-  <framebuffer>
-    <attachment usage="color" link="coords-texture-1" />
-    <use program="primary-coords-program">
-      <instance of="cube-1-group" />
-    </use>
-  </framebuffer>
-  <framebuffer>
-    <attachment usage="color" link="results-texture" />
-    <use program="first-pass-program">
-      <group id="intersection-group">
-        <booleanAnd of="cube-1 cube-2" />
-      </group>
-    </use>
-  </framebuffer>
-\end{Verbatim}
+\begin{enumerate}
+  \item Cull front faces
+  \item Bind framebuffer with coordinates texture
+  \item Activate shader program for rendering texture coordinates
+  \item Draw cube behind intersection
+  \item Bind framebuffer with results texture
+  \item Activate shader program for rendering a volume
+  \item Draw Boolean AND of cubes
+\end{enumerate}
 
@@
 
-\begin{Verbatim}[fontsize=\small]
-  <cull faces="front" />
-  <framebuffer>
-    <attachment usage="color" link="coords-texture-1" />
-    <use program="primary-coords-program">
-      <instance of="intersection-group" />
-    </use>
-  </framebuffer>
-  <framebuffer>
-    <attachment usage="color" link="coords-texture-2" />
-    <use program="secondary-coords-program">
-      <instance of="intersection-group" />
-    </use>
-  </framebuffer>
-  <cull faces="back" />
-  <framebuffer>
-    <attachment usage="color" link="third-pass-texture" />
-    <use program="third-pass-program">
-      <uniform name="CoordsTexture1" link="coords-texture-1" />
-      <uniform name="CoordsTexture2" link="coords-texture-2" />
-      <instance of="intersection-group" />
-    </use>
-  </framebuffer>
-\end{Verbatim}
+\begin{enumerate}
+  \item Cull front faces
+  \item Bind framebuffer with first coordinates texture
+  \item Activate shader program for rendering primary texture coordinates
+  \item Draw Boolean AND of cubes
+  \item Bind framebuffer with second coordinates texture
+  \item Activate shader program for rendering secondary texture coordinates
+  \item Draw Boolean AND of cubes
+  \item Cull back faces
+  \item Bind framebuffer with results texture
+  \item Activate shader program for rendering volume
+  \item Draw Boolean AND of cubes
+\end{enumerate}
 
 After that the part of the volume in front of the intersection is rendered to
-the results texture.  We have the shader program set the depth of the fragments
-generated to the lowest value to effectively mask off that part of the screen.
+the results texture.
 
-\begin{Verbatim}[fontsize=\small]
-  <cull mode="back" />
-  <framebuffer>
-    <attachment usage="color" link="coords-texture-1" />
-    <use program="primary-coords-program">
-      <uniform name="MVPMatrix" usage="modelviewprojection" />
-      <instance of="cube-1-group" />
-    </use>
-  </framebuffer>
-  <framebuffer>
-    <attachment usage="color" link="results-texture" />
-    <use program="fourth-pass-program">
-      <instance of="intersection-group" />
-    </use>
-  </framebuffer>
-\end{Verbatim}
+\begin{enumerate}
+  \item Cull back faces
+  \item Bind framebuffer with coordinates texture
+  \item Activate shader program for rendering primary texture coordinates
+  \item Draw cube in front of intersection
+  \item Bind framebuffer with results texture
+  \item Activate shader program for rendering volume
+  \item Draw Boolean AND of cubes
+  \item Mask off depth buffer
+\end{enumerate}
+
+The area of the screen with the intersection should be masked off so that
+nothing else in that area will be rendered.
+
+\begin{enumerate}
+  \item Activate shader program for setting depth to zero
+  \item Draw Boolean AND of cubes
+\end{enumerate}
 
@@
 
-\begin{Verbatim}[fontsize=\small]
-  <texture id="volume-texture-1" file="volume-1.vlb" />
-  <texture id="volume-texture-2" file="volume-2.vlb" />
-  <texture id="coords-texture-1" size="768" />
-  <texture id="coords-texture-2" size="768" />
-  <texture id="results-texture" size="768" />
-\end{Verbatim}
+\begin{enumerate}
+  \item Bind volume textures
+  \item Bind textures to store texture coordinates
+  \item Bind texture to store results
+\end{enumerate}
 
@@
 
-\begin{Verbatim}[fontsize=\small]
-  <renderbuffer id="depth-renderbuffer" format="depth" size="768" />
-\end{Verbatim}
+\begin{enumerate}
+  \item Create depth renderbuffer
+\end{enumerate}
 
@@
 
-\begin{Verbatim}[fontsize=\small]
-  <framebuffer>
-    <attachment usage="color" link="results-texture" />
-    <clear color="0 0 0 0" />
-  </framebuffer>
-\end{Verbatim}
+\begin{enumerate}
+  \item Bind framebuffer with results texture
+  \item Clear color
+\end{enumerate}
 
@@
 to \emph{LESS}.  That way we render using the faces that are farthest forward.
-Note in the code snippet below we only show the first volume.
 
-\begin{Verbatim}[fontsize=\small]
-  <framebuffer>
-    <attachment usage="color" link="coords-texture-1" />
-    <attachment usage="depth" link="depth-renderbuffer" />
-    <clear depth="0.0" />
-    <depth function="greater" />
-    <cull faces="front" />
-    <use program="primary-coords-program">
-      <group id="pieces-of-cube-1-behind-intersection-group">
-        <booleanXor of="cube-1 cube-2" hide="cube-2" filter="gequal" />
-      </group>
-    </use>
-  </framebuffer>
-  <framebuffer>
-    <attachment usage="color" link="results-texture" />
-    <attachment usage="depth" link="depth-renderbuffer" />
-    <clear depth="1.0" />
-    <depth function="less" />
-    <cull faces="back" />
-    <use program="first-pass-program">
-      <uniform name="MVPMatrix" usage="viewprojection" />
-      <uniform name="BackFacesTexture" link="coords-texture-1" />
-      <uniform name="VolumeTexture" link="volume-texture-1" />
-      <instance of="pieces-of-cube-1-behind-intersection-group" />
-    </use>
-  </framebuffer>
-\end{Verbatim}
+\begin{enumerate}
+  \item Bind framebuffer with coordinates texture and depth renderbuffer
+  \item Clear depth to zero
+  \item Set depth function to \emph{GREATER}
+  \item Cull front faces
+  \item Activate shader program for rendering primary texture coordinates
+  \item Draw Boolean XOR with pieces behind intersection
+  \item Bind framebuffer with results texture and depth renderbuffer
+  \item Clear depth to one
+  \item Set depth function to \emph{LESS}
+  \item Cull back faces
+  \item Activate shader program for rendering volumes
+  \item Draw Boolean XOR with pieces behind intersection
+\end{enumerate}
 
@@
 
-\begin{Verbatim}[fontsize=\small]
-  <cull faces="front" />
-  <framebuffer>
-    <attachment usage="color" link="coords-texture-1" />
-    <use program="primary-coords-program">
-      <uniform name="MVPMatrix" usage="viewprojection" />
-      <group id="intersection-group">
-        <booleanAnd of="cube-1 cube-2" />
-      </group>
-    </use>
-  </framebuffer>
-  <framebuffer>
-    <attachment usage="color" link="coords-texture-2" />
-    <use program="secondary-coords-program">
-      <uniform name="MVPMatrix" usage="viewprojection" />
-      <instance of="intersection-group" />
-    </use>
-  </framebuffer>
-  <cull faces="back" />
-  <framebuffer>
-    <attachment usage="color" link="results-texture" />
-    <use program="second-pass-program">
-      <uniform name="MVPMatrix" usage="viewprojection" />
-      <uniform name="ResultsTexture" link="results-texture" />
-      <uniform name="CoordsTexture1" link="coords-texture-1" />
-      <uniform name="CoordsTexture2" link="coords-texture-2" />
-      <uniofrm name="VolumeTexture1" link="volume-texture-1" />
-      <uniofrm name="VolumeTexture2" link="volume-texture-2" />
-      <instance of="intersection-group" />
-    </use>
-  </framebuffer>
-\end{Verbatim}
+\begin{enumerate}
+  \item Cull front faces
+  \item Bind framebuffer with first coordinates texture
+  \item Activate shader program for rendering primary texture coordinates
+  \item Draw Boolean AND of cubes
+  \item Bind framebuffer with second coordinates texture
+  \item Activate shader program for rendering secondary texture coordinates
+  \item Draw Boolean AND of cubes
+  \item Cull back faces
+  \item Bind framebuffer with results texture
+  \item Activate shader program for rendering volumes
+  \item Draw Boolean AND of cubes
+\end{enumerate}
 
-Next we render any pieces of the first cube in front of the intersection to the
+Next we render any pieces of the cube in front of the intersection to the
 results texture.  Similar to the first pass, we use the depth buffer to make
-sure we only take the texture coordinates of the faces we want.  Again we omit
-the code to render the second cube.
+sure we only take the texture coordinates of the faces we want.
 
-\begin{Verbatim}[fontsize=\small]
-  <framebuffer>
-    <attachment usage="color" link="coords-texture-1" />
-    <attachment usage="depth" link="depth-renderbuffer" />
-    <cull faces="front" />
-    <depth function="greater" />
-    <clear depth="0.0" />
-    <use program="primary-coords-program">
-      <uniform name="MVPMatrix" usage="viewprojection" />
-      <group id="pieces-of-cube-1-in-front-of-intersection">
-        <booleanXor of="cube-1 cube-2" hide="cube-2" filter="less" />
-      </group>
-    </use>
-  </framebuffer>
-  <framebuffer>
-    <attachment usage="color" link="results-texture" />
-    <attachment usage="depth" link="depth-renderbuffer" />
-    <depth function="less" />
-    <cull faces="back" />
-    <use program="third-pass-program">
-      <uniform name="MVPMatrix" usage="viewprojection" />
-      <uniform name="ResultsTexture" link="results-texture" />
-      <uniform name="CoordsTexture" link="coords-texture-1" />
-      <uniform name="VolumeTexture" link="volume-texture-1" />
-      <insance of="pieces-of-cube-1-in-front-of-intersection" />
-    </use>
-  </framebuffer>
-\end{Verbatim}
+\begin{enumerate}
+  \item Bind framebuffer with coordinates texture and depth renderbuffer
+  \item Cull front faces
+  \item Set depth function to \emph{GREATER}
+  \item Clear depth buffer to zero
+  \item Activate shader program for rendering primary texture coordinates
+  \item Draw Boolean XOR with pieces in front of intersection
+  \item Bind framebuffer with results texture and depth renderbuffer
+  \item Cull back faces
+  \item Set depth function to \emph{LESS}
+  \item Activate shader program for rendering volumes
+  \item Draw Boolean XOR with pieces in front of intersection
+\end{enumerate}
 
@@
 
-\begin{Verbatim}[fontsize=\small]
-  <use program="final-pass-program">
-    <uniform name="ResultsTexture" link="results-texture" />
-    <translate x="-1.0">
-      <uniform name="MVPMatrix" usage="modelviewprojection" />
-      <cube id="cube-1" />
-    </translate>
-    <translate x="+1.0">
-      <uniform name="MVPMatrix" usage="modelviewprojection" />
-      <cube id="cube-2" />
-    </translate>
-  </use>
-\end{Verbatim}
+\begin{enumerate}
+  \item Activate shader for rendering results
+  \item Render both cubes
+\end{enumerate}
 

### 2013-04-22 paper.tex
Only one cube in front of intersection in second approach
In figure for first approach, mask off with 'Boolean AND' not 'intersection'

--- 2013-04-22 paper.tex
@@
 intersection, colored blue, is rendered.  Finally the depth buffer is masked off
-using the intersection, and everything else, colored yellow, is rendered.
+using the Boolean AND, and everything else, colored yellow, is rendered.
 }

### 2013-04-22 paper.tex
Better explanation of yellow parts in caption for first approach figure
Split up discussion of libraries in implementation section
Say what SVG means
Mention implementation will work with OpenGL 4 too

--- 2013-04-22 paper.tex
@@
 of OpenGL 3.  In other words, no deprecated functionality from OpenGL 2 was
-used.
+used.  Furthermore, because OpenGL 4 is a superset of the OpenGL 3 core profile,
+the implementation will easily work with that as well.
 
 To facilitate the process, a handful of general-purpose graphics libraries were
-developed.  At the lowest levels are a simple 3D mathematics library and a
-library providing thin C++ wrappers around the components already in OpenGL.
-Above that, another library provides additional utilities for working with
-OpenGL, including one for loading volumetric data as a texture.  At the highest
-level is a library providing a simple render graph that can be used to rapidly
-prototype OpenGL applications.  Similar to SVG, the render graph can be written
-in XML so scenes can easily be saved, reloaded, and shared.  All of the
-libraries can be compiled as shared libraries using GCC on Linux and Mac OS X.
+developed.
+
+At the lowest levels are a simple 3D mathematics library and a library providing
+thin C++ wrappers around the components already in OpenGL.  Above that, another
+library provides additional utilities for working with OpenGL, including one for
+loading volumetric data as a texture.
+
+At the highest level is a library providing a simple render graph that can be
+used to rapidly prototype OpenGL applications.  Similar to SVG, or Scalable
+Vector Graphics, the render graph can be written in XML so scenes can easily be
+saved, reloaded, and shared.
+
+All of the libraries can be compiled as shared libraries using GCC on Linux and
+Mac OS X.
 

### 2013-04-22 paper.tex
Mention not having to recompile application with RapidGL
Mention name of RapidGL

@@
 
-At the highest level is a library providing a simple render graph that can be
-used to rapidly prototype OpenGL applications.  Similar to SVG, or Scalable
-Vector Graphics, the render graph can be written in XML so scenes can easily be
-saved, reloaded, and shared.
+At the highest level is a library named {\em RapidGL} providing a simple render
+graph that can be used to rapidly prototype OpenGL applications.  Similar to
+SVG, or Scalable Vector Graphics, the render graph can be written in XML so
+scenes can easily be saved, reloaded, and shared.  In addition, users are free
+to experiment with their scenes without ever recompiling the application.
 

### 2013-04-22 paper.tex
Explain Node class in RapidGL

@@
 
+Core to RapidGL is the {\tt Node} class.  Nodes can override the {\tt visit}
+method to perform actions when the scene is traversed.  Most implementations
+perform any setup which requires other nodes the first time they are visited.
+For example, many nodes check for ancestors with specific types or search the
+scene for nodes with a specific ID.  Nodes can also register themselves as {\em
+listeners} of other nodes by implementing the {\tt NodeListener} interface and
+passing a pointer to the node's {\tt addNodeListener} method.
+
 All of the libraries can be compiled as shared libraries using GCC on Linux and

### 2013-04-22 paper.tex
Show example XML code, use to explain instancing

@@
 
+The XML below shows how the second pass of the first approach is rendered.
+Notice that the render graph supports basic instancing through the {\em group}
+and {\em instance} nodes.  When an instance node is visited, it simply traverses
+the group it points to.  Complex behavior can be achieved with nested groups.
+
+\begin{Verbatim}[fontsize=\small]
+  <cull faces="front" />
+  <framebuffer>
+    <attachment usage="color" link="coords-texture-1" />
+    <use program="primary-coords-program">
+      <group id="intersection-group">
+        <booleanAnd of="cube-1 cube-2" />
+      </group>
+    </use>
+  </framebuffer>
+  <framebuffer>
+    <attachment usage="color" link="coords-texture-2" />
+    <use program="secondary-coords-program">
+      <instance of="intersection-group" />
+    </use>
+  </framebuffer>
+  <cull faces="back" />
+  <framebuffer>
+    <attachment usage="color" link="results-texture" />
+    <use program="second-pass-program">
+      <uniform name="CoordsTexture1" link="coords-texture-1" />
+      <uniform name="CoordsTexture2" link="coords-texture-2" />
+      <uniform name="VolumeTexture1" link="volume-texture-1" />
+      <uniform name="VolumeTexture2" link="volume-texture-2" />
+      <uniform name="MVPMatrix" usage="modelviewprojection" />
+      <instance of="intersection-group" />
+    </use>
+  </framebuffer>
+\end{Verbatim}
+
 All of the libraries can be compiled as shared libraries using GCC on Linux and

### 2013-04-22 paper.tex
Explain how Boolean nodes were implemented

@@
 
+The Boolean operations required by the two approaches were implemented as two
+separate nodes.  Currently both nodes take two cubes, which are assumed to be
+axis-aligned.  {\tt BooleanAndNode} finds the intersection of the two cubes by
+finding the minimums and maximums of the cubes in world space.  {\tt
+BooleanXorNode} does the same, but then uses a variation of the well-known
+Sutherland-Hodgman polyline-clipping algorithm to break up the cubes into pieces
+around the intersection.  Both nodes use the {\tt NodeListener} interface to
+only update their geometry when the cubes move.
+
 All of the libraries can be compiled as shared libraries using GCC on Linux and

### 2013-04-22 paper.tex
Removed note about some scene graph components in application
Say application views RapidGL scenes
Second sentence of application paragraph has 'users' as subject

@@
 
-From there an application was created to view the resulting scenes.  It can
-rotate the camera around the scene, as well as select objects in the scene and
-drag them around.  It shows the current frame rate so we can evaluate the
-performance of the scenes.  Some scene graph components that are not general
-enough to be in the library were created at the application level.
+From there an application was created to view RapidGL scenes.  Users can
+rotate the camera around a scene, as well as select objects and drag them.  It
+shows the current frame rate so we can evaluate the performance of the scenes.
 

### 2013-04-22 paper.tex
Clarified last sentence of paragraph about application
Fixed order of 'they' in questions in evaluation section
Note some libraries require Poco, GLFW, and/or CppUnit

@@
 All of the libraries can be compiled as shared libraries using GCC on Linux and
-Mac OS X.
+Mac OS X.  Some libraries require {\em Poco} for text and XML utilities.  {\em
+GLFW} and {\em CppUnit} are also required by some libraries for unit testing.
 
 From there an application was created to view RapidGL scenes.  Users can
-rotate the camera around a scene, as well as select objects and drag them.  It
-shows the current frame rate so we can evaluate the performance of the scenes.
+rotate the camera around a scene, as well as select objects and drag them.  The
+application shows the current frame rate so the performance of a scene can be
+evaluated.
 
@@
 
-How fast they can be rendered?
-How easy they were to implement?
+How fast can they be rendered?
+How easy were they to implement?
 

### 2013-04-22 paper.tex
Mention Glycerin includes a text renderer

@@
 library provides additional utilities for working with OpenGL, including one for
-loading volumetric data as a texture.
+loading volumetric data as a texture, and another for rendering text.
 

### 2013-04-22 paper.tex
Use unordered list for questions in evaluation section
Added screenshots
Figure for second approach says it's top down

@@
 \caption{
-Rendering intersecting volumes using the Boolean AND with Boolean XOR approach.
-First the pieces behind the intersection, colored red, are rendered.  Next the
-intersection, colored green, is rendered.  Then the pieces in front of the
-intersection, colored blue, are rendered.  A final pass renders the results to
-the screen.
+Top down view of rendering intersecting volumes using the Boolean AND with
+Boolean XOR approach.  First the pieces behind the intersection, colored red,
+are rendered.  Next the intersection, colored green, is rendered.  Then the
+pieces in front of the intersection, colored blue, are rendered.  A final pass
+renders the results to the screen.
 }
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=0.95\textwidth]{boolean-and-screenshot-1.png}
+\caption{Screenshot of Boolean AND.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.95\textwidth]{boolean-xor-screenshot-1.png}
+\caption{Screenshot of Boolean XOR.}
+\end{figure}
+
 % ==========
@@
 
-How fast can they be rendered?
-How easy were they to implement?
+\begin{itemize}
+  \item How fast can they be rendered?
+  \item How easy were they to implement?
+\end{itemize}
 

### 2013-04-22 paper.tex
Fixed 'top-down' not being hyphenated

@@
 \caption{
-Top down view of rendering intersecting volumes using the Boolean AND with depth
+Top-down view of rendering intersecting volumes using the Boolean AND with depth
 buffer masking approach.  First the part of the volume directly behind the
@@
 \caption{
-Top down view of rendering intersecting volumes using the Boolean AND with
+Top-down view of rendering intersecting volumes using the Boolean AND with
 Boolean XOR approach.  First the pieces behind the intersection, colored red,

### 2013-04-22 paper.tex
Elude to slicing method needing to generate a large amount of geometry every frame in abstract

@@
 rendering an arbitrary number of intersecting volumes on the GPU have been
-developed, but they can require a significant amount of effort to implement.
-This paper investigates implementing the simplest case, where two volumes
-overlap, using familiar boolean operations and a traditional graphics library.
+developed, but they can require a large amount of geometry to be created every
+frame, or may be difficult or time-consuming to implement.  This paper
+investigates implementing the simplest case, where two volumes overlap, using
+familiar boolean operations and a traditional graphics library.
 \end{abstract}

### 2013-04-22 paper.tex
Better statement of purpose in abstract, mentions limited to axis-aligned volumes

@@
 frame, or may be difficult or time-consuming to implement.  This paper
-investigates implementing the simplest case, where two volumes overlap, using
-familiar boolean operations and a traditional graphics library.
+investigates whether using familiar Boolean operations may be faster or easy to
+implement in the simple case where two axis-aligned volumes intersect.
 \end{abstract}

### 2013-04-22 paper.tex
Abstract says 'problematic' instead of 'difficult'

@@
 \begin{abstract}
-Rendering intersecting volumes is difficult because the renderer must alternate
-between sampling from each volume to achieve the correct blending.  Methods for
-rendering an arbitrary number of intersecting volumes on the GPU have been
-developed, but they can require a large amount of geometry to be created every
-frame, or may be difficult or time-consuming to implement.  This paper
-investigates whether using familiar Boolean operations may be faster or easy to
-implement in the simple case where two axis-aligned volumes intersect.
+Rendering intersecting volumes is problematic because the renderer must
+alternate between sampling from each volume to achieve the correct blending.
+Methods for rendering an arbitrary number of intersecting volumes on the GPU
+have been developed, but they can require a large amount of geometry to be
+created every frame, or may be difficult or time-consuming to implement.  This
+paper investigates whether using familiar Boolean operations may be faster or
+easy to implement in the simple case where two axis-aligned volumes intersect.
 \end{abstract}

### 2013-05-26 paper.tex
Added new screenshots (ref #9)
Added basic table with some performance measurements (ref #4)
Frame rate for slicing on Macbook (ref #4)
Frame rate for optimized slicing on Macbook (ref #4)
Removed bullets in Evaluation section (ref #14)
Initial evaluation section (ref #14)

@@
 \setlength{\pdfpageheight}{11in}
+\usepackage{booktabs}
 \usepackage{graphicx}
@@
 \centering
-\includegraphics[width=0.95\textwidth]{boolean-and-screenshot-1.png}
+\includegraphics[width=0.95\textwidth]{boolean-and-screenshot-2.png}
 \caption{Screenshot of Boolean AND.}
@@
 \centering
-\includegraphics[width=0.95\textwidth]{boolean-xor-screenshot-1.png}
+\includegraphics[width=0.95\textwidth]{boolean-xor-screenshot-2.png}
 \caption{Screenshot of Boolean XOR.}
@@
 
-\begin{itemize}
-  \item How fast can they be rendered?
-  \item How easy were they to implement?
-\end{itemize}
+The two approaches to rendering intersecting volumes using boolean operations
+were evaluated using different criteria.
+
+Because volume visualization is used in the scientific and medical communities
+to aid in making important decisions, an approach that does not handle all cases
+will likely not be appropriate for widespread usage.  With that in mind, the
+accuracy of each approach was judged by looking for edge cases with visual
+artifacts.
+
+Unfortunately the Boolean XOR approach may show some visual artifacts from
+certain angles.  This occurs because when two pieces have the same depth value,
+due to perspective one of the pieces may actually appear behind the other.  In
+the future, this issue may be solved by using the distance from the camera to
+sort rather than the actual depth.
+
+% Picture of visual artifacts with Boolean XOR approach
+
+% Picture of depth-sorting issue
+
+The performance of each approach was measured by using a simple frame counter
+showing the number of frames rendered in the previous second.  Multiple camera
+angles and scene configurations were tried in order to expose strong and weak
+points of the approaches.  The tests were run on at least three different
+machines with a variety of operating systems (Linux, Mac OS X) and video cards
+(ATI, NVIDIA).  Table \ref{performance-table} shows the results.
+
+\begin{table}
+  \centering
+  \begin{tabular}{ l l l }
+    \toprule
+    Approach & Macbook & Desktop \\
+    \midrule
+    Boolean AND with masking & 22 & 13 \\
+    Boolean AND with XOR & 24 & 13 \\
+    Slicing & 6 & ? \\
+    Optimized Slicing & 25 & ? \\
+    \bottomrule
+  \end{tabular}
+  \caption{Performance measurements}
+  \label{performance-table}
+\end{table}
+
+% Furthermore, a variety of sampling rates were used to determine how each
+% approach scales as the number of samples taken increases.
+
+As a baseline the Boolean-based approaches were compared to the existing
+view-aligned slicing method from previous work by developing a simple prototype.
+Since the slicing method requires significantly more state changes when the
+volumes are side-by-side as opposed to when one is in front of the other, it was
+particularly important with this method that performance measurements were taken
+from a variety of perspectives.
+
+In addition, a variation of the slicing method that was optimized for two
+volumes was created.  Basically rather than using uniform variables to switch
+textures as would normally be done, the optimized variation stores the texture
+with each vertex instead.  While it requires a fairly significant amount of
+extra storage, using vertex attributes allows all the slices to be transferred
+to the card at once.  The result is a considerable performance boost, similar to
+how it is generally faster to transfer one large file across a network then many
+small ones.
+
+As shown both Boolean-based approaches outperformed the traditional slicing
+method, but were in turn bested by the optimized slicing method.  It should be
+noted that the frame rates for the optimized slicing method are the worst case,
+and in average it performed even better, around forty frames per second.
+
+Judging from the numbers, it appears that the multiple passes of the
+Boolean-based approaches create too much overhead compared to the optimized
+slicing method.  The optimized slicing method also has a very simple fragment
+shader which may keep it from being fragment-bound.
+
+Altogether, although the Boolean-based approaches outperform the un-optimized
+slicing method, it is recommended that the traditional slicing method with the
+optimizations described be used.  The optimized slicing method is not only
+faster, but it supports volumes of any orientation.  Furthermore, the optimized
+method should scale to more volumes relatively easily.
 

### 2013-05-26 paper.tex
Added screenshot of visual artifact with Boolean XOR (ref #14)
Mention figure of artifact right after bringing it up (ref #14)
Figure showing problem with pieces in perspective (ref #14)

@@
 Unfortunately the Boolean XOR approach may show some visual artifacts from
-certain angles.  This occurs because when two pieces have the same depth value,
-due to perspective one of the pieces may actually appear behind the other.  In
-the future, this issue may be solved by using the distance from the camera to
-sort rather than the actual depth.
+certain angles.  Figure \ref{boolean-xor-with-artifact} shows an example of the
+artifact.  This occurs because when two pieces have the same depth value, due to
+perspective one of the pieces may actually appear behind the other, as
+illustrated in figure \ref{pieces-in-perspective}.  In the future, this issue
+may be solved by using the distance from the camera to sort rather than the
+actual depth.
 
-% Picture of visual artifacts with Boolean XOR approach
+\begin{figure}
+\centering
+\includegraphics[width=0.95\textwidth]{boolean-xor-screenshot-3.png}
+\caption{Screenshot of Boolean XOR with visual artifact.}
+\label{boolean-xor-with-artifact}
+\end{figure}
 
-% Picture of depth-sorting issue
+\begin{figure}
+\centering
+\includegraphics[width=0.95\textwidth]{pieces-in-perspective.pdf}
+\caption{Top-down view of two pieces in perspective.  Notice how even though
+both pieces have the same depth, a small sliver of the blue piece on the left
+ends up being rendered behind the red piece on the right.}
+\label{pieces-in-perspective}
+\end{figure}
 

### 2013-05-30 paper.tex
Blurb about how lessons learned may still be useful to future work (ref #14)
Include chart for frame rates from gnuplot
Added chart of frame rates using PGFPlots

@@
 \usepackage{fancyvrb}
+\usepackage{pgfplots}
 
@@
 
+\begin{figure}
+\centering
+\includegraphics{frame-rates.pdf}
+\label{frame-rates}
+\end{figure}
+
+\begin{figure}
+\centering
+\begin{tikzpicture}
+\begin{axis}[
+    xlabel=Samples,
+    ylabel=Frames Per Second]
+\addplot coordinates {
+  (100, 22)
+  (200, 10)
+  (300, 7)
+  (400, 5)
+  (500, 4)
+};
+\addlegendentry{Slicing (attributes)}
+\addplot coordinates {
+  (100, 20)
+  (200, 10)
+  (300, 7)
+  (400, 5)
+  (500, 4)
+};
+\addlegendentry{Boolean XOR}
+\addplot coordinates {
+  (100, 20)
+  (200, 10)
+  (300, 6)
+  (400, 4)
+  (500, 4)
+};
+\addlegendentry{Depth-masking}
+\addplot coordinates {
+  (100, 6)
+  (200, 3)
+  (300, 2)
+  (400, 1)
+  (500, 1)
+};
+\addlegendentry{Slicing (uniforms)}
+\end{axis}
+\end{tikzpicture}
+\caption{
+Chart showing performance measurements as number of samples varies.  The slicing
+with attributes, boolean XOR, and depth-masking methods are all about the same,
+while the slicing with uniforms method lags behind considerably.
+}
+\end{figure}
+
 Altogether, although the Boolean-based approaches outperform the un-optimized
@@
 method should scale to more volumes relatively easily.
+
+However, it is hoped that the lessons learned in this paper, namely that the
+overhead incurred by multi-pass algorithms may outweigh the cost of streaming
+geometry in other algorithms, will prove valuable to future endeavors in both
+volume rendering and graphics in general.
 

### 2013-06-02 paper.tex
Removed gnuplot-based frame rates chart
Added a title to the performance chart
Rearranged datasets in performance chart
Added chart for performance results from desktop

@@
 \centering
-\includegraphics{frame-rates.pdf}
-\label{frame-rates}
-\end{figure}
-
-\begin{figure}
-\centering
 \begin{tikzpicture}
 \begin{axis}[
+    title=2.2 GHz i7 with AMD Radeon HD 6750M,
     xlabel=Samples,
@@
 \addplot coordinates {
-  (100, 22)
+  (100, 20)
   (200, 10)
-  (300, 7)
-  (400, 5)
+  (300, 6)
+  (400, 4)
   (500, 4)
 };
-\addlegendentry{Slicing (attributes)}
+\addlegendentry{Depth-masking}
 \addplot coordinates {
@@
 \addplot coordinates {
-  (100, 20)
-  (200, 10)
-  (300, 6)
-  (400, 4)
-  (500, 4)
-};
-\addlegendentry{Depth-masking}
-\addplot coordinates {
   (100, 6)
@@
 \addlegendentry{Slicing (uniforms)}
+\addplot coordinates {
+  (100, 22)
+  (200, 10)
+  (300, 7)
+  (400, 5)
+  (500, 4)
+};
+\addlegendentry{Slicing (attributes)}
 \end{axis}
@@
 while the slicing with uniforms method lags behind considerably.
+}
+\end{figure}
+
+\begin{figure}
+\centering
+\begin{tikzpicture}
+\begin{axis}[
+    title=3.0 GHz P4 with NVIDIA GeForce 8500 GT,
+    xlabel=Samples,
+    ylabel=Frames Per Second]
+\addplot coordinates {
+  (100, 10)
+  (200, 6)
+  (300, 4)
+  (400, 3)
+  (500, 2)
+};
+\addlegendentry{Depth-masking}
+\addplot coordinates {
+  (100, 13)
+  (200, 7)
+  (300, 5)
+  (400, 4)
+  (500, 3)
+};
+\addlegendentry{Boolean XOR}
+\addplot coordinates {
+  (100, 18)
+  (200, 9)
+  (300, 6)
+  (400, 5)
+  (500, 4)
+};
+\addlegendentry{Slicing (uniforms)}
+\addplot coordinates {
+  (100, 15)
+  (200, 8)
+  (300, 5)
+  (400, 4)
+  (500, 3)
+};
+\addlegendentry{Slicing (attributes)}
+\end{axis}
+\end{tikzpicture}
+\caption{
+Chart showing performance measurements as number of samples varies.  Here the
+slicing with uniforms method is fastest, followed by the slicing with attributes
+method, and the Boolean XOR method.  The depth-masking method is slowest.
 }

### 2013-06-02 paper.tex
Abstract should say 'easier' not 'easy'

@@
 paper investigates whether using familiar Boolean operations may be faster or
-easy to implement in the simple case where two axis-aligned volumes intersect.
+easier to implement in the simple case where two axis-aligned volumes intersect.
 \end{abstract}

### 2013-06-02 paper.tex
Say 'using hardware acceleration' instead of 'on the GPU' in abstract

@@
 alternate between sampling from each volume to achieve the correct blending.
-Methods for rendering an arbitrary number of intersecting volumes on the GPU
-have been developed, but they can require a large amount of geometry to be
-created every frame, or may be difficult or time-consuming to implement.  This
-paper investigates whether using familiar Boolean operations may be faster or
-easier to implement in the simple case where two axis-aligned volumes intersect.
+Methods for rendering an arbitrary number of intersecting volumes using hardware
+acceleration have been developed, but they can require a large amount of
+geometry to be created every frame, or may be difficult or time-consuming to
+implement.  This paper investigates whether using familiar Boolean operations
+may be faster or easier to implement in the simple case where two axis-aligned
+volumes intersect.
 \end{abstract}

### 2013-06-03 paper.tex
Say 'generally require' instead of 'can require' in abstract

--- 2013-06-02 paper.tex
@@
 alternate between sampling from each volume to achieve the correct blending.
-Methods for rendering an arbitrary number of intersecting volumes on the GPU
-have been developed, but they can require a large amount of geometry to be
-created every frame, or may be difficult or time-consuming to implement.  This
-paper investigates whether using familiar Boolean operations may be faster or
-easier to implement in the simple case where two axis-aligned volumes intersect.
+Methods for rendering an arbitrary number of intersecting volumes using hardware
+acceleration have been developed, but they generally require a large amount of
+geometry to be created every frame, or may be difficult or time-consuming to
+implement.  This paper investigates whether using familiar Boolean operations
+may be faster or easier to implement in the simple case where two axis-aligned
+volumes intersect.
 \end{abstract}

### 2013-06-03 paper.tex
Don't need 'familiar' in abstract

@@
 geometry to be created every frame, or may be difficult or time-consuming to
-implement.  This paper investigates whether using familiar Boolean operations
-may be faster or easier to implement in the simple case where two axis-aligned
-volumes intersect.
+implement.  This paper investigates whether using Boolean operations may be
+faster or easier to implement in the simple case where two axis-aligned volumes
+intersect.
 \end{abstract}

### 2013-06-03 paper.tex
Abstract doesn't talk about how easy/difficult to implement

@@
 acceleration have been developed, but they generally require a large amount of
-geometry to be created every frame, or may be difficult or time-consuming to
-implement.  This paper investigates whether using Boolean operations may be
-faster or easier to implement in the simple case where two axis-aligned volumes
-intersect.
+geometry to be created every frame.  This paper investigates whether using
+Boolean operations may be faster in the simple case where two axis-aligned
+volumes intersect.
 \end{abstract}

### 2013-06-03 paper.tex
Abstract says 'are faster' instead of 'may be faster'
Abstract sums up results

@@
 geometry to be created every frame.  This paper investigates whether using
-Boolean operations may be faster in the simple case where two axis-aligned
-volumes intersect.
+Boolean operations are faster in the simple case where two axis-aligned volumes
+intersect.  Although viable options it is shown that two new methods are only as
+fast or no faster than the traditional slicing method.
 \end{abstract}

### 2013-06-03 paper.tex
Added slicing measurements for desktop
Added labels for performance charts so they can be referenced
Reference performance charts in text

@@
     Boolean AND with XOR & 24 & 13 \\
-    Slicing & 6 & ? \\
-    Optimized Slicing & 25 & ? \\
+    Slicing & 6 & 18 \\
+    Optimized Slicing & 25 & 15 \\
     \bottomrule
@@
 
-% Furthermore, a variety of sampling rates were used to determine how each
-% approach scales as the number of samples taken increases.
+Furthermore, a variety of sampling rates were used to determine how each
+approach scales as the number of samples taken increases.  Figure
+\ref{macbook-performance-chart} shows how the methods scale on a 2.2 GHz Intel
+Core i7 Macbook Pro with an AMD Radeon HD 6750M video card.  Likewise figure
+\ref{desktop-performance-chart} does so for an older 3.0 GHz Intel Pentium 4
+Linux desktop with an NVIDIA GeForce 8500 GT video card.
 
@@
 }
+\label{macbook-performance-chart}
 \end{figure}
@@
 }
+\label{desktop-performance-chart}
 \end{figure}

### 2013-06-03 paper.tex
Capitalized 'Accelerated' in title
Just say 'no faster than' in abstract

@@
 % =====
-\title{Hardware-accelerated Rendering of Intersecting Volumes using Boolean Operations}
+\title{Hardware-Accelerated Rendering of Intersecting Volumes using Boolean Operations}
 \author{Andrew Brown \and Joe Geigel}
@@
 Boolean operations are faster in the simple case where two axis-aligned volumes
-intersect.  Although viable options it is shown that two new methods are only as
-fast or no faster than the traditional slicing method.
+intersect.  Although viable options it is shown that two new methods are no
+faster than the traditional slicing method.
 \end{abstract}

### 2013-06-03 paper.tex
Abstract points out geometry doesn't need to be regenerated if only camera is moving
Abstract explicitly says frame rates rather than just talking about speed

@@
 acceleration have been developed, but they generally require a large amount of
-geometry to be created every frame.  This paper investigates whether using
-Boolean operations are faster in the simple case where two axis-aligned volumes
-intersect.  Although viable options it is shown that two new methods are no
-faster than the traditional slicing method.
+geometry to be created every frame.  In the simple case where two axis-aligned
+volumes intersect, it is shown that Boolean operations can be used so that
+the camera may move without regenerating any geometry.  However, the two
+variations presented herein still achieve similar frame rates to that of the
+traditional slicing method.
 \end{abstract}

### 2013-06-03 paper.tex
Abstract says 'canonical' instead of 'traditional'
Only tested on two machines so far

@@
 variations presented herein still achieve similar frame rates to that of the
-traditional slicing method.
+canonical slicing method.
 \end{abstract}
@@
 angles and scene configurations were tried in order to expose strong and weak
-points of the approaches.  The tests were run on at least three different
-machines with a variety of operating systems (Linux, Mac OS X) and video cards
-(ATI, NVIDIA).  Table \ref{performance-table} shows the results.
+points of the approaches.  The tests were run on two different machines with a
+variety of operating systems (Linux, Mac OS X) and video cards (ATI, NVIDIA).
+Table \ref{performance-table} shows the results.
 

### 2013-06-04 paper.tex
Added results from workstation

@@
 angles and scene configurations were tried in order to expose strong and weak
-points of the approaches.  The tests were run on two different machines with a
+points of the approaches.  The tests were run on three different machines with a
 variety of operating systems (Linux, Mac OS X) and video cards (ATI, NVIDIA).
@@
   \centering
-  \begin{tabular}{ l l l }
+  \begin{tabular}{ l l l l }
     \toprule
-    Approach & Macbook & Desktop \\
+    Approach & Macbook & Desktop & Workstation\\
     \midrule
-    Boolean AND with masking & 22 & 13 \\
-    Boolean AND with XOR & 24 & 13 \\
-    Slicing & 6 & 18 \\
-    Optimized Slicing & 25 & 15 \\
+    Boolean AND with masking & 22 & 13 & 51 \\
+    Boolean AND with XOR & 24 & 13 & 76 \\
+    Slicing & 6 & 18 & 130 \\
+    Optimized Slicing & 25 & 15 & 172 \\
     \bottomrule
@@
 \ref{desktop-performance-chart} does so for an older 3.0 GHz Intel Pentium 4
-Linux desktop with an NVIDIA GeForce 8500 GT video card.
+Linux desktop with an NVIDIA GeForce 8500 GT video card.  Furthermore, figure
+\ref{workstation-performance-chart} depicts the results for a newer 3.33 GHz
+Intel Core i7 Linux workstation with an NVIDIA Quadro FX 5800 video card.
 
@@
 
+\begin{figure}
+\centering
+\begin{tikzpicture}
+\begin{axis}[
+    title=3.33 GHz Intel Core i7 with Quadro FX 5800,
+    xlabel=Samples,
+    ylabel=Frames Per Second]
+\addplot coordinates {
+  (100, 51)
+  (200, 28)
+  (300, 19)
+  (400, 14)
+  (500, 11)
+};
+\addlegendentry{Depth-masking}
+\addplot coordinates {
+  (100, 76)
+  (200, 39)
+  (300, 27)
+  (400, 23)
+  (500, 17)
+};
+\addlegendentry{Boolean XOR}
+\addplot coordinates {
+  (100, 130)
+  (200, 68)
+  (300, 46)
+  (400, 36)
+  (500, 29)
+};
+\addlegendentry{Slicing (uniforms)}
+\addplot coordinates {
+  (100, 172)
+  (200, 89)
+  (300, 60)
+  (400, 46)
+  (500, 38)
+};
+\addlegendentry{Slicing (attributes)}
+\end{axis}
+\end{tikzpicture}
+\caption{
+Chart showing performance measurements as number of samples varies on an Intel
+Core i7 system with a Quadro FX 5800 video card.  Here the difference between
+the methods is more dramatic.  Both slicing methods handedly beat the
+Boolean-based methods.
+}
+\label{workstation-performance-chart}
+\end{figure}
+
 Altogether, although the Boolean-based approaches outperform the un-optimized

### 2013-06-13 paper.tex
Right-align numbers in performance table
Fixed some typos

@@
 
-For those reasons, in this paper we examine techniques for rending two volumes
+For those reasons, in this paper we examine techniques for rendering two volumes
 using simple boolean operations.  Although limited, we hope the techniques
@@
 called fragments.  The fragments are passed along to the last stage, which
-shades the them, decides which ones to discard, and displays those that remain
-as pixels.
+shades them, decides which ones to discard, and displays those that remain as
+pixels.
 
@@
 fragment generated by rendering the bounding boxes creates its own ray.  The
-rays start at the viewpoint and passes through the fragment’s location.  Any
+rays start at the viewpoint and pass through the fragment’s location.  Any
 volumes that intersect the ray are sampled at regular intervals along it by
@@
   \centering
-  \begin{tabular}{ l l l l }
+  \begin{tabular}{ l r r r }
     \toprule
@@
 noted that the frame rates for the optimized slicing method are the worst case,
-and in average it performed even better, around forty frames per second.
+and on average it performed even better, around forty frames per second.
 

### 2013-06-20 paper.tex
Don't use empty page style for first page
Use report instead of article
Use adapted suthesis-2e.sty
Replaced sections with chapters
Replaced subsections with sections
Revised abstract for new results

@@
-\documentclass{article}
+\documentclass{report}
+\usepackage{suthesis-2e}
 \setlength{\pdfpagewidth}{8.5in}
@@
 
-% Use an empty page style for the title and abstract
-\pagestyle{empty}
-
 % =====
@@
 \title{Hardware-Accelerated Rendering of Intersecting Volumes using Boolean Operations}
-\author{Andrew Brown \and Joe Geigel}
-\date{\small{Rochester Institute of Technology}}
-\maketitle
+\author{Andrew Brown}
+\dept{Computer Science}
+\committeechair{Professor Joe Geigel, Ph.D.}
+\reader{Professor Reynold Bailey, Ph.D.}
+\observer{Professor Warren Carithers}
+\grantdate{June}{2013}
+
+\beforepreface
+\afterpreface
 
@@
 volumes intersect, it is shown that Boolean operations can be used so that
-the camera may move without regenerating any geometry.  However, the two
-variations presented herein still achieve similar frame rates to that of the
-canonical slicing method.
+the camera may move without regenerating any geometry.  Furthermore, samples
+that do not contribute to the final image may be completely avoided.  The two
+variations presented herin achieve higher frame rates than the traditional
+slicing method.
 \end{abstract}
@@
 % ============
-\section{Introduction}
+\chapter{Introduction}
 
@@
 % ==========
-\section{Background}
+\chapter{Background}
 
@@
 % =============
-\section{Previous Work}
+\chapter{Previous Work}
 
-\subsection{Overview}
+\section{Overview}
 
@@
 
-\subsection{Slicing}
+\section{Slicing}
 
@@
 
-\section{Ray Casting}
+\chapter{Ray Casting}
 
@@
 % ==========
-\section{Approaches}
+\chapter{Approaches}
 
@@
 
-\subsection{Boolean AND with Depth Buffer Masking}
+\section{Boolean AND with Depth Buffer Masking}
 
@@
 
-\subsection{Boolean AND with Boolean XOR}
+\section{Boolean AND with Boolean XOR}
 
@@
 % ==============
-\section{Implementation}
+\chapter{Implementation}
 
@@
 % ==========
-\section{Evaluation}
+\chapter{Evaluation}
 

### 2013-06-20 paper.tex
Replaced screenshots
Replaced results in charts

@@
 \centering
-\includegraphics[width=0.95\textwidth]{boolean-and-screenshot-2.png}
+\includegraphics[width=0.95\textwidth]{boolean-and-screenshot.png}
 \caption{Screenshot of Boolean AND.}
@@
 \centering
-\includegraphics[width=0.95\textwidth]{boolean-xor-screenshot-2.png}
+\includegraphics[width=0.95\textwidth]{boolean-xor-screenshot.png}
 \caption{Screenshot of Boolean XOR.}
@@
 \begin{axis}[
-    title=2.2 GHz i7 with AMD Radeon HD 6750M,
+    title=2.2 GHz Intel Core i7 with AMD Radeon HD 6750M,
     xlabel=Samples,
@@
 \addplot coordinates {
-  (100, 20)
-  (200, 10)
+  (100, 12)
+  (200, 9)
   (300, 6)
-  (400, 4)
-  (500, 4)
-};
-\addlegendentry{Depth-masking}
-\addplot coordinates {
-  (100, 20)
-  (200, 10)
-  (300, 7)
   (400, 5)
@@
 \addplot coordinates {
-  (100, 6)
-  (200, 3)
-  (300, 2)
+  (100, 12)
+  (200, 9)
+  (300, 6)
+  (400, 5)
+  (500, 4)
+};
+\addlegendentry{Depth-Masking}
+\addplot coordinates {
+  (100, 3)
+  (200, 1)
+  (300, 1)
   (400, 1)
@@
 };
-\addlegendentry{Slicing (uniforms)}
+\addlegendentry{General Slicing}
 \addplot coordinates {
-  (100, 22)
-  (200, 10)
-  (300, 7)
-  (400, 5)
-  (500, 4)
+  (100, 12)
+  (200, 6)
+  (300, 4)
+  (400, 3)
+  (500, 2)
 };
-\addlegendentry{Slicing (attributes)}
+\addlegendentry{Optimized Slicing}
 \end{axis}
@@
 \caption{
-Chart showing performance measurements as number of samples varies.  The slicing
-with attributes, boolean XOR, and depth-masking methods are all about the same,
-while the slicing with uniforms method lags behind considerably.
+Chart showing performance measurements as number of samples varies on an Intel
+Core i7 system with an AMD Radeon HD 6750M video card.
 }
@@
 \addplot coordinates {
-  (100, 51)
-  (200, 28)
-  (300, 19)
+  (100, 38)
+  (200, 24)
+  (300, 18)
   (400, 14)
-  (500, 11)
-};
-\addlegendentry{Depth-masking}
-\addplot coordinates {
-  (100, 76)
-  (200, 39)
-  (300, 27)
-  (400, 23)
-  (500, 17)
+  (500, 12)
 };
@@
 \addplot coordinates {
-  (100, 130)
-  (200, 68)
-  (300, 46)
-  (400, 36)
-  (500, 29)
+  (100, 37)
+  (200, 25)
+  (300, 18)
+  (400, 15)
+  (500, 12)
 };
-\addlegendentry{Slicing (uniforms)}
+\addlegendentry{Depth-Masking}
 \addplot coordinates {
-  (100, 172)
-  (200, 89)
-  (300, 60)
-  (400, 46)
-  (500, 38)
+  (100, 32)
+  (200, 17)
+  (300, 11)
+  (400, 9)
+  (500, 7)
 };
-\addlegendentry{Slicing (attributes)}
+\addlegendentry{General Slicing}
+\addplot coordinates {
+  (100, 35)
+  (200, 18)
+  (300, 12)
+  (400, 9)
+  (500, 7)
+};
+\addlegendentry{Optimized Slicing}
 \end{axis}
@@
 Chart showing performance measurements as number of samples varies on an Intel
-Core i7 system with a Quadro FX 5800 video card.  Here the difference between
-the methods is more dramatic.  Both slicing methods handedly beat the
-Boolean-based methods.
+Core i7 system with a Quadro FX 5800 video card.
 }

### 2013-06-23 paper.tex
Use listing for showing XML (ref #15)

@@
 \usepackage{graphicx}
+\usepackage[T1]{fontenc}
 \usepackage[utf8]{inputenc}
 \usepackage{fancyvrb}
+\usepackage{listings}
 \usepackage{pgfplots}
@@
 
-The XML below shows how the second pass of the first approach is rendered.
-Notice that the render graph supports basic instancing through the {\em group}
-and {\em instance} nodes.  When an instance node is visited, it simply traverses
-the group it points to.  Complex behavior can be achieved with nested groups.
-
-\begin{Verbatim}[fontsize=\small]
+The XML in listing \ref{listing:xml-of-second-pass} shows how the second pass of
+the first approach is rendered.  Notice that the render graph supports basic
+instancing through the {\em group} and {\em instance} nodes.  When an instance
+node is visited, it simply traverses the group it points to.  Complex behavior
+can be achieved with nested groups.
+
+\lstset{language=XML}
+\lstset{markfirstintag=true}
+\lstset{showstringspaces=false}
+\lstset{caption=XML of second pass for first approach.}
+\lstset{label=listing:xml-of-second-pass}
+\begin{lstlisting}[frame=single,float]
 <cull faces="front" />
@@
 </framebuffer>
-\end{Verbatim}
+\end{lstlisting}
 

### 2013-06-23 paper.tex
Don't indent all of XML listing
Fixed misspelling of 'herein' in abstract (ref #18)
Abstract focuses more on slicing method (ref #18)

@@
 alternate between sampling from each volume to achieve the correct blending.
-Methods for rendering an arbitrary number of intersecting volumes using hardware
-acceleration have been developed, but they generally require a large amount of
-geometry to be created every frame.  In the simple case where two axis-aligned
-volumes intersect, it is shown that Boolean operations can be used so that
-the camera may move without regenerating any geometry.  Furthermore, samples
-that do not contribute to the final image may be completely avoided.  The two
-variations presented herin achieve higher frame rates than the traditional
-slicing method.
+Unfortunately the traditional method for rendering intersecting volumes using
+hardware acceleration requires a large amount of geometry to be created every
+frame.  In the simple case where two axis-aligned volumes intersect, it is shown
+that Boolean operations can be used so that the camera may move without
+regenerating any geometry.  Furthermore, samples that do not contribute to the
+final image may be completely avoided.  The two variations presented herein
+achieve higher frame rates than the traditional slicing method.
 \end{abstract}

### 2013-06-23 paper.tex
Emphasize early ray termination in abstract (ref #18)

@@
 that Boolean operations can be used so that the camera may move without
-regenerating any geometry.  Furthermore, samples that do not contribute to the
-final image may be completely avoided.  The two variations presented herein
-achieve higher frame rates than the traditional slicing method.
+regenerating any geometry.  Originally the two variations presented herein
+still achieved similar frame rates.  However, after implementing early ray
+termination, they have proven faster for typical data sets.
 \end{abstract}

### 2013-06-23 paper.tex
Sorted Boolean XOR method before Depth-Masking in Desktop performance chart (ref #19)
Updated results in desktop performance chart (ref #19)
Updated title of desktop performance chart (ref #19)
Updated caption of desktop performance chart (ref #19)

@@
 \begin{axis}[
-    title=3.0 GHz P4 with NVIDIA GeForce 8500 GT,
+    title=3.0 GHz Intel Pentium 4 with NVIDIA GeForce 8500 GT,
     xlabel=Samples,
@@
 \addplot coordinates {
-  (100, 10)
-  (200, 6)
-  (300, 4)
-  (400, 3)
-  (500, 2)
+  (100, 4)
+  (200, 3)
+  (300, 2)
+  (400, 1)
+  (500, 1)
+};
+\addlegendentry{Boolean XOR}
+\addplot coordinates {
+  (100, 4)
+  (200, 2)
+  (300, 2)
+  (400, 1)
+  (500, 1)
 };
@@
 \addplot coordinates {
-  (100, 13)
-  (200, 7)
-  (300, 5)
-  (400, 4)
-  (500, 3)
-};
-\addlegendentry{Boolean XOR}
-\addplot coordinates {
-  (100, 18)
-  (200, 9)
-  (300, 6)
-  (400, 5)
-  (500, 4)
+  (100, 3)
+  (200, 1)
+  (300, 1)
+  (400, 1)
+  (500, 1)
 };
@@
 \addplot coordinates {
-  (100, 15)
-  (200, 8)
-  (300, 5)
-  (400, 4)
-  (500, 3)
+  (100, 3)
+  (200, 1)
+  (300, 1)
+  (400, 1)
+  (500, 1)
 };
@@
 \caption{
-Chart showing performance measurements as number of samples varies.  Here the
-slicing with uniforms method is fastest, followed by the slicing with attributes
-method, and the Boolean XOR method.  The depth-masking method is slowest.
+Chart showing performance measurements as number of samples varies on an Intel
+Pentium 4 system with an NVIDIA GeForce 8500 GT video card.
 }

### 2013-06-23 paper.tex
Removed sentence about multiple camera angles (ref #17)
Updated numbers in performance table (ref #17)

@@
 The performance of each approach was measured by using a simple frame counter
-showing the number of frames rendered in the previous second.  Multiple camera
-angles and scene configurations were tried in order to expose strong and weak
-points of the approaches.  The tests were run on three different machines with a
-variety of operating systems (Linux, Mac OS X) and video cards (ATI, NVIDIA).
-Table \ref{performance-table} shows the results.
+showing the number of frames rendered in the previous second.  The tests were
+run on three different machines with a variety of operating systems (Linux, Mac
+OS X) and video cards (ATI, NVIDIA).  Table \ref{performance-table} shows the
+results.
 
@@
     \midrule
-    Boolean AND with masking & 22 & 13 & 51 \\
-    Boolean AND with XOR & 24 & 13 & 76 \\
-    Slicing & 6 & 18 & 130 \\
-    Optimized Slicing & 25 & 15 & 172 \\
+    Boolean AND with masking & 12 & 4 & 37 \\
+    Boolean AND with XOR & 12 & 4 & 38 \\
+    Slicing & 12 & 3 & 32 \\
+    Optimized Slicing & 3 & 3 & 35 \\
     \bottomrule

### 2013-06-23 paper.tex
Sorted rows in performance table (ref #17)
Fixed names of approaches in table and charts (ref #17)
Removed blurb about stage changes with slicing method (ref #17)

@@
     \midrule
-    Boolean AND with masking & 12 & 4 & 37 \\
-    Boolean AND with XOR & 12 & 4 & 38 \\
-    Slicing & 12 & 3 & 32 \\
+    Boolean XOR & 12 & 4 & 38 \\
+    Depth-Masking & 12 & 4 & 37 \\
+    General Slicing & 12 & 3 & 32 \\
     Optimized Slicing & 3 & 3 & 35 \\
@@
 view-aligned slicing method from previous work by developing a simple prototype.
-Since the slicing method requires significantly more state changes when the
-volumes are side-by-side as opposed to when one is in front of the other, it was
-particularly important with this method that performance measurements were taken
-from a variety of perspectives.
 
@@
 };
-\addlegendentry{Depth-masking}
+\addlegendentry{Depth-Masking}
 \addplot coordinates {
@@
 };
-\addlegendentry{Slicing (uniforms)}
+\addlegendentry{General Slicing}
 \addplot coordinates {
@@
 };
-\addlegendentry{Slicing (attributes)}
+\addlegendentry{Optimized Slicing}
 \end{axis}

### 2013-06-23 paper.tex
Consolidated paragraphs about slicing method variants (ref #17)
First pass at redoing evaluation prose (ref #17)

@@
 view-aligned slicing method from previous work by developing a simple prototype.
-
-In addition, a variation of the slicing method that was optimized for two
-volumes was created.  Basically rather than using uniform variables to switch
-textures as would normally be done, the optimized variation stores the texture
-with each vertex instead.  While it requires a fairly significant amount of
-extra storage, using vertex attributes allows all the slices to be transferred
-to the card at once.  The result is a considerable performance boost, similar to
-how it is generally faster to transfer one large file across a network then many
-small ones.
-
-As shown both Boolean-based approaches outperformed the traditional slicing
-method, but were in turn bested by the optimized slicing method.  It should be
-noted that the frame rates for the optimized slicing method are the worst case,
-and on average it performed even better, around forty frames per second.
-
-Judging from the numbers, it appears that the multiple passes of the
-Boolean-based approaches create too much overhead compared to the optimized
-slicing method.  The optimized slicing method also has a very simple fragment
-shader which may keep it from being fragment-bound.
+To be fair, a variation of the slicing method optimized for two volumes was
+created.  Basically rather than using uniform variables to switch textures as
+would normally be done, the optimized variation stores the texture with each
+vertex instead.  While it requires a fairly significant amount of extra storage,
+using vertex attributes allows all the slices to be transferred to the card at
+once.  The result is a considerable performance boost, similar to how it is
+generally faster to transfer one large file across a network then many small
+ones.
+
+As shown the Boolean-based approaches generally outperformed the two variants of
+the slicing method.  The difference between the frame rates becomes more clear
+as the number of samples increases.
+
+Judging from the numbers, it appears all the approaches are fragment-bound, and
+thus the Boolean-based approaches win out because they can employ early ray
+termination.
 
@@
 
-Altogether, although the Boolean-based approaches outperform the un-optimized
-slicing method, it is recommended that the traditional slicing method with the
-optimizations described be used.  The optimized slicing method is not only
-faster, but it supports volumes of any orientation.  Furthermore, the optimized
-method should scale to more volumes relatively easily.
-
-However, it is hoped that the lessons learned in this paper, namely that the
-overhead incurred by multi-pass algorithms may outweigh the cost of streaming
-geometry in other algorithms, will prove valuable to future endeavors in both
-volume rendering and graphics in general.
+Altogether, it is difficult to say for sure which approach developers should
+take to render intersecting volumes.  If the application only needs to handle
+two axis-aligned volumes, using one of the Boolean-based approaches could
+provide a more interactive experience for the user.  However, if the application
+needs more than two volumes, or if those volumes need to be oriented
+arbitrarily, it is best to stick with the slicing method.
+
+In the future it would be interesting to see if either of the two Boolean-based
+approaches could be expanded to support more volumes and different orientations.
+The Boolean XOR approach would most likely be easier to extend to more volumes,
+since it is just a matter of creating more pieces.  However, the depth-masking
+approach will be better off for supporting different orientations because it
+works in pixel space.
+
+Either way, it is hoped that the lessons learned in this report, namely that
+rendering volumes is inherently fragment-bound and avoiding any unnecessary
+processing with early ray termination is key to achieving significant speed-ups,
+will prove valuable to future future endeavors in both volume rendering and
+graphics in general.
 

### 2013-06-24 paper.tex
Venn diagrams for Boolean operations
Just say depth-masking instead of depth-buffer masking

@@
 two shapes, while XOR, meaning exclusive OR, takes everything but the
-intersection.
+intersection.  Figures \ref{boolean-and-venn-diagram} and
+\ref{boolean-xor-venn-diagram} show the Boolean AND and Boolean XOR operations
+as Venn diagrams.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\textwidth]{venn-and.pdf}
+\caption{Boolean AND.}
+\label{boolean-and-venn-diagram}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\textwidth]{venn-xor.pdf}
+\caption{Boolean XOR.}
+\label{boolean-xor-venn-diagram}
+\end{figure}
 
@@
 
-\section{Boolean AND with Depth Buffer Masking}
+\section{Boolean AND with Depth-Masking}
 
@@
 \caption{
-Top-down view of rendering intersecting volumes using the Boolean AND with depth
-buffer masking approach.  First the part of the volume directly behind the
+Top-down view of rendering intersecting volumes using the Boolean AND with
+depth-masking approach.  First the part of the volume directly behind the
 intersection, colored red, is rendered.  Next the intersection, colored green,

### 2013-06-24 paper.tex
Fixed a missing 'to' in Depth-Masking approach section

@@
 The first approach uses a Boolean AND operation combined with depth-buffer
-masking.  The basic idea is render the part of the volume directly behind the
+masking.  The basic idea is to render the part of the volume directly behind the
 intersection, render the intersection itself, and then render the part of the

### 2013-06-24 paper.tex
Removed paragraphs about approaches being CPU- or GPU-based

@@
 
-It is mostly GPU-based, only requiring that the Boolean operation be performed
-on the CPU.  While there is relatively less work on the CPU, it does require a
-significant amount of passes.
-
 The first step in this approach is to bind all the textures.  There are two for
@@
 \end{enumerate}
-
-It is mostly CPU-based, since it requires both Boolean operations to be done on
-the CPU.  The extra work on the CPU results in less rendering passes, however.
-The pseudocode for the approach is shown below.
 

### 2013-06-24 paper.tex
Don't need 'current' when talking about why OpenGL 3
Diagram of clipping process for Boolean XOR

@@
 The two approaches were implemented using OpenGL with C++.  To ensure results
-are relevant to current and future applications, we used only the core profile
-of OpenGL 3.  In other words, no deprecated functionality from OpenGL 2 was
-used.  Furthermore, because OpenGL 4 is a superset of the OpenGL 3 core profile,
-the implementation will easily work with that as well.
+are relevant to future applications, we used only the core profile of OpenGL 3.
+In other words, no deprecated functionality from OpenGL 2 was used.
+Furthermore, because OpenGL 4 is a superset of the OpenGL 3 core profile, the
+implementation will easily work with that as well.
 
@@
 Sutherland-Hodgman polyline-clipping algorithm to break up the cubes into pieces
-around the intersection.  Both nodes use the {\tt NodeListener} interface to
-only update their geometry when the cubes move.
+around the intersection, as shown in figure \ref{clipping}.  Both nodes use the
+{\tt NodeListener} interface to only update their geometry when the cubes move.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\textwidth]{clipping.pdf}
+\caption{Variation of Sutherland-Hodgman clipping algorithm in Boolean XOR
+implementation.}
+\label{clipping}
+\end{figure}
 

### 2013-06-24 paper.tex
Fixed caption of depth-masking screenshot
Say 'approach' in screenshot captions for approaches
Say 'on some systems' when talking about performance boost of optimized slicing

@@
 \includegraphics[width=0.95\textwidth]{boolean-and-screenshot.png}
-\caption{Screenshot of Boolean AND.}
+\caption{Screenshot of Depth-Masking approach.}
 \end{figure}
@@
 \includegraphics[width=0.95\textwidth]{boolean-xor-screenshot.png}
-\caption{Screenshot of Boolean XOR.}
+\caption{Screenshot of Boolean XOR approach.}
 \end{figure}
@@
 using vertex attributes allows all the slices to be transferred to the card at
-once.  The result is a considerable performance boost, similar to how it is
-generally faster to transfer one large file across a network then many small
-ones.
+once.  The result is a considerable performance boost on some systems, similar
+to how it is generally faster to transfer one large file across a network then
+many small ones.
 

### 2013-06-24 paper.tex
Removed duplicate 'future' in last paragraph

@@
 processing with early ray termination is key to achieving significant speed-ups,
-will prove valuable to future future endeavors in both volume rendering and
-graphics in general.
+will prove valuable to future endeavors in both volume rendering and graphics in
+general.
 

### 2013-06-24 paper.tex
Put caption of listing below it
Diagram for volume representations
Listing for Boolean AND implementation

@@
 individual values are named voxels, similar to how the elements of a picture are
-called pixels.
+called pixels.  Figure \ref{representations} shows two different representations
+of a volume.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.9\textwidth]{representations.pdf}
+\caption{Representations of a volume.  One can think of a volume as a 3D grid of
+voxels or as a stack of 2D images.}
+\label{representations}
+\end{figure}
 
@@
 \lstset{caption=XML of second pass for first approach.}
+\lstset{captionpos=b}
 \lstset{label=listing:xml-of-second-pass}
@@
 axis-aligned.  {\tt BooleanAndNode} finds the intersection of the two cubes by
-finding the minimums and maximums of the cubes in world space.  {\tt
-BooleanXorNode} does the same, but then uses a variation of the well-known
-Sutherland-Hodgman polyline-clipping algorithm to break up the cubes into pieces
-around the intersection, as shown in figure \ref{clipping}.  Both nodes use the
-{\tt NodeListener} interface to only update their geometry when the cubes move.
+finding the minimums and maximums of the cubes in world space, as shown in
+listing \ref{listing:boolean-and}.  {\tt BooleanXorNode} does the same, but then
+uses a variation of the well-known Sutherland-Hodgman polyline-clipping
+algorithm to break up the cubes into pieces around the intersection, as shown in
+figure \ref{clipping}.  Both nodes use the {\tt NodeListener} interface to only
+update their geometry when the cubes move.
+
+\lstset{language=C++}
+\lstset{caption=Boolean AND implementation.}
+\lstset{captionpos=b}
+\lstset{label=listing:boolean-and}
+\begin{lstlisting}[frame=single,float]
+Vec4 min = max(c1.min, c2.min);
+Vec4 max = min(c1.max, c2.max);
+\end{lstlisting}
 

### 2013-06-24 paper.tex
Replaced performance table with chart

@@
 run on three different machines with a variety of operating systems (Linux, Mac
-OS X) and video cards (ATI, NVIDIA).  Table \ref{performance-table} shows the
-results.
-
-\begin{table}
-  \centering
-  \begin{tabular}{ l r r r }
-    \toprule
-    Approach & Macbook & Desktop & Workstation\\
-    \midrule
-    Boolean XOR & 12 & 4 & 38 \\
-    Depth-Masking & 12 & 4 & 37 \\
-    General Slicing & 12 & 3 & 32 \\
-    Optimized Slicing & 3 & 3 & 35 \\
-    \bottomrule
-  \end{tabular}
-  \caption{Performance measurements}
-  \label{performance-table}
-\end{table}
+OS X) and video cards (ATI, NVIDIA).  Figure 7.3 shows the results with one
+hundred samples.
+
+\begin{figure}
+\label{overall-performance-chart}
+\centering
+\begin{tikzpicture}
+\begin{axis}[
+    xbar,
+    xmin=0,
+    xlabel=Frames Per Second,
+    enlarge y limits=0.2,
+    symbolic y coords={Optimized Slicing, General Slicing, Depth-Masking, Boolean XOR},
+    ytick=data
+]
+\addplot coordinates {
+    (2,General Slicing)
+    (12,Optimized Slicing)
+    (12,Depth-Masking)
+    (12,Boolean XOR)
+};
+\addplot coordinates {
+    (32,General Slicing)
+    (35,Optimized Slicing)
+    (37,Depth-Masking)
+    (38,Boolean XOR)
+};
+\addplot coordinates {
+    (3,General Slicing)
+    (3,Optimized Slicing)
+    (4,Depth-Masking)
+    (4,Boolean XOR)
+};
+\end{axis}
+\end{tikzpicture}
+\caption{Performance of approaches on three different systems with 100 samples.}
+\end{figure}
 

### 2013-06-24 paper.tex
Clean up last sentence of abstract
Clean up motiviation in introduction

@@
 still achieved similar frame rates.  However, after implementing early ray
-termination, they have proven faster for typical data sets.
+termination, they have proven to be faster for typical data sets.
 \end{abstract}
@@
 different methods have been developed for rendering an arbitrary number of
-volumes using hardware acceleration, but they can be problematic to implement,
-rely on general-purpose GPU programming, or both.
+volumes using hardware acceleration, but they can be inefficient, problematic to
+implement, or both.
 

### 2013-06-24 paper.tex
Switched order of approaches

@@
 
-\section{Boolean AND with Depth-Masking}
-
-The first approach uses a Boolean AND operation combined with depth-buffer
-masking.  The basic idea is to render the part of the volume directly behind the
-intersection, render the intersection itself, and then render the part of the
-volume directly in front of the intersection.  Finally the area of the
-intersection is masked off and everything else is drawn depth-sorted with
-blending enabled.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.8\textwidth]{boolean-and.pdf}
-\caption{
-Top-down view of rendering intersecting volumes using the Boolean AND with
-depth-masking approach.  First the part of the volume directly behind the
-intersection, colored red, is rendered.  Next the intersection, colored green,
-is rendered.  Then the the part of the volume directly in front of the
-intersection, colored blue, is rendered.  Finally the depth buffer is masked off
-using the Boolean AND so that when the cubes are rendered again, only the parts
-colored yellow are rendered.
-}
-\label{boolean-and}
-\end{figure}
-
-The first step in this approach is to bind all the textures.  There are two for
-the volumes, two for storing the texture coordinates of the back faces of the
-cubes, and one for storing results between passes.  The last three textures will
-be used with external framebuffers, so they should have a width and height at
-least as big as the window.
-
-\begin{enumerate}
-  \item Bind volume textures
-  \item Bind textures to store texture coordinates
-  \item Bind texture to store results
-\end{enumerate}
-
-Next the part of the volume behind the intersection is rendered into the results
-texture.  To get just that part the back faces of the Boolean AND are used
-instead of the front faces of the cube.
-
-\begin{enumerate}
-  \item Cull front faces
-  \item Bind framebuffer with coordinates texture
-  \item Activate shader program for rendering texture coordinates
-  \item Draw cube behind intersection
-  \item Bind framebuffer with results texture
-  \item Activate shader program for rendering a volume
-  \item Draw Boolean AND of cubes
-\end{enumerate}
-
-From there the intersection itself is rendered.  Because the intersection needs
-to alternate between samples from both volumes, we have to render the back face
-coordinates into two different textures, and then make them both available to
-the shader program that renders the intersection.
-
-\begin{enumerate}
-  \item Cull front faces
-  \item Bind framebuffer with first coordinates texture
-  \item Activate shader program for rendering primary texture coordinates
-  \item Draw Boolean AND of cubes
-  \item Bind framebuffer with second coordinates texture
-  \item Activate shader program for rendering secondary texture coordinates
-  \item Draw Boolean AND of cubes
-  \item Cull back faces
-  \item Bind framebuffer with results texture
-  \item Activate shader program for rendering volume
-  \item Draw Boolean AND of cubes
-\end{enumerate}
-
-After that the part of the volume in front of the intersection is rendered to
-the results texture.
-
-\begin{enumerate}
-  \item Cull back faces
-  \item Bind framebuffer with coordinates texture
-  \item Activate shader program for rendering primary texture coordinates
-  \item Draw cube in front of intersection
-  \item Bind framebuffer with results texture
-  \item Activate shader program for rendering volume
-  \item Draw Boolean AND of cubes
-  \item Mask off depth buffer
-\end{enumerate}
-
-The area of the screen with the intersection should be masked off so that
-nothing else in that area will be rendered.
-
-\begin{enumerate}
-  \item Activate shader program for setting depth to zero
-  \item Draw Boolean AND of cubes
-\end{enumerate}
-
-Finally we render the two volumes again depth-sorted with blending enabled and
-the depth function set to \emph{LESS}.  Since on the previous step we masked off
-the area of the screen with the intersection, the fragments in those parts will
-just be discarded.  In fact, on most modern video cards they will not even be
-processed by the fragment shader at all because of early Z-kill.
-
-\begin{enumerate}
-  \item Enable blending
-  \item Set depth function to \emph{LESS}
-  \item Bind framebuffer with coordinates texture
-  \item Cull front faces
-  \item Activate shader program for rendering primary texture coordinates
-  \item Draw cube behind intersection
-  \item Cull back faces
-  \item Activate shader program for rendering volumes
-  \item Draw cube behind intersection
-  \item Repeat for cube in front of intersection
-\end{enumerate}
-
 \section{Boolean AND with Boolean XOR}
 
-The second approach uses a Boolean AND operation together with a Boolean XOR.
+The first approach uses a Boolean AND operation together with a Boolean XOR.
 Here the idea is to break up both volumes into pieces around the intersection.
@@
 
-Like the first approach, the first step is to set up the textures.  There are
-two textures for the volumes, two to store the texture coordinates of the back
+The first step in this approach is to set up the textures.  There are two
+textures for the volumes, two to store the texture coordinates of the back
 faces, and one to hold results between passes.  The textures to store the
@@
   \item Render both cubes
+\end{enumerate}
+
+\section{Boolean AND with Depth-Masking}
+
+The second approach uses a Boolean AND operation combined with depth-buffer
+masking.  The basic idea is to render the part of the volume directly behind the
+intersection, render the intersection itself, and then render the part of the
+volume directly in front of the intersection.  Finally the area of the
+intersection is masked off and everything else is drawn depth-sorted with
+blending enabled.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.8\textwidth]{boolean-and.pdf}
+\caption{
+Top-down view of rendering intersecting volumes using the Boolean AND with
+depth-masking approach.  First the part of the volume directly behind the
+intersection, colored red, is rendered.  Next the intersection, colored green,
+is rendered.  Then the the part of the volume directly in front of the
+intersection, colored blue, is rendered.  Finally the depth buffer is masked off
+using the Boolean AND so that when the cubes are rendered again, only the parts
+colored yellow are rendered.
+}
+\label{boolean-and}
+\end{figure}
+
+Like the first approach, the first step is to bind all the textures.  There are
+two for the volumes, two for storing the texture coordinates of the back faces
+of the cubes, and one for storing results between passes.  The last three
+textures will be used with external framebuffers, so they should have a width
+and height at least as big as the window.
+
+\begin{enumerate}
+  \item Bind volume textures
+  \item Bind textures to store texture coordinates
+  \item Bind texture to store results
+\end{enumerate}
+
+Next the part of the volume behind the intersection is rendered into the results
+texture.  To get just that part the back faces of the Boolean AND are used
+instead of the front faces of the cube.
+
+\begin{enumerate}
+  \item Cull front faces
+  \item Bind framebuffer with coordinates texture
+  \item Activate shader program for rendering texture coordinates
+  \item Draw cube behind intersection
+  \item Bind framebuffer with results texture
+  \item Activate shader program for rendering a volume
+  \item Draw Boolean AND of cubes
+\end{enumerate}
+
+From there the intersection itself is rendered.  Because the intersection needs
+to alternate between samples from both volumes, we have to render the back face
+coordinates into two different textures, and then make them both available to
+the shader program that renders the intersection.
+
+\begin{enumerate}
+  \item Cull front faces
+  \item Bind framebuffer with first coordinates texture
+  \item Activate shader program for rendering primary texture coordinates
+  \item Draw Boolean AND of cubes
+  \item Bind framebuffer with second coordinates texture
+  \item Activate shader program for rendering secondary texture coordinates
+  \item Draw Boolean AND of cubes
+  \item Cull back faces
+  \item Bind framebuffer with results texture
+  \item Activate shader program for rendering volume
+  \item Draw Boolean AND of cubes
+\end{enumerate}
+
+After that the part of the volume in front of the intersection is rendered to
+the results texture.
+
+\begin{enumerate}
+  \item Cull back faces
+  \item Bind framebuffer with coordinates texture
+  \item Activate shader program for rendering primary texture coordinates
+  \item Draw cube in front of intersection
+  \item Bind framebuffer with results texture
+  \item Activate shader program for rendering volume
+  \item Draw Boolean AND of cubes
+  \item Mask off depth buffer
+\end{enumerate}
+
+The area of the screen with the intersection should be masked off so that
+nothing else in that area will be rendered.
+
+\begin{enumerate}
+  \item Activate shader program for setting depth to zero
+  \item Draw Boolean AND of cubes
+\end{enumerate}
+
+Finally we render the two volumes again depth-sorted with blending enabled and
+the depth function set to \emph{LESS}.  Since on the previous step we masked off
+the area of the screen with the intersection, the fragments in those parts will
+just be discarded.  In fact, on most modern video cards they will not even be
+processed by the fragment shader at all because of early Z-kill.
+
+\begin{enumerate}
+  \item Enable blending
+  \item Set depth function to \emph{LESS}
+  \item Bind framebuffer with coordinates texture
+  \item Cull front faces
+  \item Activate shader program for rendering primary texture coordinates
+  \item Draw cube behind intersection
+  \item Cull back faces
+  \item Activate shader program for rendering volumes
+  \item Draw cube behind intersection
+  \item Repeat for cube in front of intersection
 \end{enumerate}
@@
 \centering
+\includegraphics[width=0.95\textwidth]{boolean-xor-screenshot.png}
+\caption{Screenshot of Boolean XOR approach.}
+\end{figure}
+
+\begin{figure}
+\centering
 \includegraphics[width=0.95\textwidth]{boolean-and-screenshot.png}
 \caption{Screenshot of Depth-Masking approach.}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.95\textwidth]{boolean-xor-screenshot.png}
-\caption{Screenshot of Boolean XOR approach.}
 \end{figure}

### 2013-06-24 paper.tex
Changed 'mentioned' to 'discussed'
Don't need to capitalize 'depth-masking' in caption
Say 'approach' in visual artifact screenshot

@@
 Both approaches are based on the volume rendering method in \cite{Kruger03}, as
-previously mentioned.  To render part of a volume, they first render the texture
+previously discussed.  To render part of a volume, they first render the texture
 coordinates of the back faces of that part to a texture.  Then the front faces
@@
 \includegraphics[width=0.95\textwidth]{boolean-and-screenshot.png}
-\caption{Screenshot of Depth-Masking approach.}
+\caption{Screenshot of depth-masking approach.}
 \end{figure}
@@
 \includegraphics[width=0.95\textwidth]{boolean-xor-screenshot-3.png}
-\caption{Screenshot of Boolean XOR with visual artifact.}
+\caption{Screenshot of Boolean XOR approach with visual artifact.}
 \label{boolean-xor-with-artifact}

### 2013-06-24 paper.tex
Screenshot for simple shader
Screenshot for texture coords
Green and red circles in fragment ray-casting diagram

@@
 illuminated.  Many times they look up values stored in textures, which are
-generally one, two, or three dimensional images.
+generally one, two, or three dimensional images.  Figure \ref{fig:xmas} shows
+the results of a simple fragment shader.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.8\textwidth]{xmas.png}
+\caption{Simple fragment shader coloring the left side of a face red and the
+right side green.}
+\label{fig:xmas}
+\end{figure}
 
@@
 0.0 to 1.0 according to the bounds of the polygon, with 0.0 being one extreme
-and 1.0 being the other.
+and 1.0 being the other.  Figure \ref{fig:coords} shows the texture coordinates
+of a cube.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.8\textwidth]{coords.png}
+\caption{Texture coordinates of a cube.}
+\label{fig:coords}
+\end{figure}
 
@@
 blended into the framebuffer at the fragment’s position.  Figure
-\ref{fragment-ray-casting} gives a depiction of the process, with green circles
-representing fragments.
+\ref{fragment-ray-casting} gives a depiction of the process, with green and red
+circles representing fragments.
 

### 2013-06-24 paper.tex
Screenshots for passes of Kruger method
Fixed XML listing for swapped order of approaches

@@
 \ref{fragment-ray-casting} gives a depiction of the process, with green and red
-circles representing fragments.
+circles representing fragments.  Figures \ref{fig:first-pass-of-kruger-method}
+and \ref{fig:second-pass-of-kruger-method} show the first and second passes of
+the Kruger method.
 
@@
 \label{fragment-ray-casting}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.8\textwidth]{back-faces.png}
+\caption{First pass of Kruger method.}
+\label{fig:first-pass-of-kruger-method}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.8\textwidth]{kruger.png}
+\caption{Second pass of Kruger method.}
+\label{fig:second-pass-of-kruger-method}
 \end{figure}
@@
 
-The XML in listing \ref{listing:xml-of-second-pass} shows how the second pass of
-the first approach is rendered.  Notice that the render graph supports basic
-instancing through the {\em group} and {\em instance} nodes.  When an instance
-node is visited, it simply traverses the group it points to.  Complex behavior
-can be achieved with nested groups.
+The XML in listing \ref{listing:xml-of-second-pass} shows how the second pass
+for the depth-masking approach is rendered.  Notice that the render graph
+supports basic instancing through the {\em group} and {\em instance} nodes.
+When an instance node is visited, it simply traverses the group it points to.
+Complex behavior can be achieved with nested groups.
 
@@
 \lstset{showstringspaces=false}
-\lstset{caption=XML of second pass for first approach.}
+\lstset{caption=XML of second pass for depth-masking approach.}
 \lstset{captionpos=b}
